### 2013-11-21 Writeups/CSS739 Paper/Paper.tex
Lots of writeups

\documentclass{article}
	\usepackage{natbib}

\begin{document}

\title{Simulating International Energy Security}
	\author{David Masad}
	\maketitle

\section{Introduction}

Energy security is defined by the International Energy Agency as ``the uninterrupted availability of energy sources at an affordable price''\citep{iea_2013}. As the world's energy demand continues to expand in the face of growing scarcity, energy security is becoming increasingly critical not just to developed countries but to developing ones as well \citep{yergin_2006}. Thus, assessing the energy security of different countries is important in order to understand the overall state of the international system. Despite this, there have been few published attempts to create formal models of energy security.

One such model, MOSES (Model of Short-term Energy Security), was created within the International Energy Agency (IEA), the intergovernmental organization of the world's major oil importers created as a counterbalance to exporter cartel OPEC \citep{}. As the name indicates, MOSES focuses on short-term risks, defined as a timescale of days and weeks. Furthermore, MOSES assigns risk scores based on an impressive, but ultimately static, set of indicators. In particular, the indicator used to measure political stability is ``weighted average of political stability of suppliers''. However, the greater the diversity of suppliers (another input into the model), the greater the potential range of political stabilities. Furthermore, averaging ratings of political risk eliminates the complexity that characterises international political stabilitiy, ignoring the possibility of crisis and conflict contagion \citep{}.

Other models of energy security focus on long-term timeframes, in years and decades. [[EXPAND MORE HERE]]

In this paper, I present a novel model linking energy security with international political stability.

\section{Model Description}

\subsection{Overview}

The model consists of countries, each of which has a set supply of and demand for crude oil. Countries are linked by a network of export-import relationships. Each month, some countries enter crisis with fixed probabilities; the exports and imports of countries in crisis are considered \emph{at risk}. If \textbf{contagion} is enabled, a crisis in one country may spread to its immediate geographic neighbors. If \textbf{assistance} is enabled, oil exporters may increase their production to balance a loss of secure oil by their major import partners.

The model parameters currently remain fixed for each run: political instability, supply, demand, and trade relations are treated as constant.

\subsection{Data Sources}
The model inputs are empirical whenever possible. Data inputs are: oil trade relationships, country political instability, and country consumption of domestic oil production. Total supply and demand are derived from input data.

The network of oil imports and exports comes from the United Nations-maintained COMTRADE database \citep{un_2013}. I extract all imports of crude petroleum (product code 2709) and other unprocessed petrolem (product code 2710) for 2012, the most recent year for which full data is available. Note that import data appears to be more reliable than export data: many major oil producers (particularly Saudi Arabia and Venezuela) only report export volume by region (e.g. Europe, North America); however, the majority of their trade partners report their imports from them.

Political instability estimates are taken from the Economist Intelligence Unit's Political Index ratings \citep(eiu_2013). These are in turn estimated based on the methodology developed by the Political Instability Task Force \citep{goldstone_2005}, which could predict the onset of internal instability in a two-year period with 80\% accuracy. Thus, these scores are normalized such that a rating of 10 (the maximum) is associated with an 80\% probability of instability within a 24-month timeframe, translating into a [[FILLIN]] monthly probability of crisis onset.

\section{Model Analysis}

\section{Discussion}

In this poster, I will present an initial implementation of a novel agent-based model linking crude oil energy security to inter- and intra-state crises and conflicts, through the mechanism of supply shocks \citep{kilian_2008}. The model operates at the meso (months-years) timescale, and allows countries to react to changes in the oil supply. It can generate ranges of notional near-future trajectories of global and local energy security, providing risk estimates for different countries. The model attempts to capture the emergent consequences of conflict contagion \citep{black_2013}, temporary production increases and varying levels of stockpiles. It can also be used to explore different scenarios, from particular crises to realignments of major trade relationships.

The agents in the model are geospatially located countries, acting in monthly timesteps. Countries experience exogenous crises based on probabilities derived from empirical estimates (such as \citep{goldstone_2005,eiu_2013,ward_2013}). They are linked geographically, and by oil trade relationships obtained via the COMTRADE database \citep{un_2013}. As countries experience crises, their participation in the  international oil system is placed at risk, changing the volume of `secure' imports or exports of other countries and  potentially triggering supply or demand shocks.

Countries respond to the changing energy security environment in several ways. Exporters may choose to temporarily increase production in response to supply shocks elsewhere, while countries that are members of the International Energy Agency may vote to release their emergency stockpiles. The model's behavior is verified against previous qualitative and quantitative analysis, and in consultation with subject-matter experts.

The poster will describe the model's methodology and input data, and provide a visual representation of the model's estimated energy security risk for every country. It will provide a visual example of a representative model run, and illustrate the results of a sweep of the model's behavior and parameter space. Finally, it will describe future work, incorporating different energy sources, shipping chokepoints, and dynamic network updating.

 \bibliographystyle{chicago}
\bibliography{energysecurity}

\end{document}

### 2013-11-21 Writeups/SBP Writeup/Poster_Writeup.tex
Lots of writeups

--- 2013-11-21 Writeups/CSS739 Paper/Paper.tex
@@
-\documentclass{article}
-	\usepackage{natbib}
+\documentclass{llncs}
 
@@
 	\author{David Masad}
+	\institute{Department of Computational Social Science \linebreak George Mason University \linebreak Fairfax, Virginia, USA}
 	\maketitle
 
-\section{Introduction}
+Energy security is defined by the International Energy Agency as ``the uninterrupted availability of energy sources at an affordable price''\cite{iea_2013}.
+As the world's energy demand continues to expand in the face of growing scarcity, energy security is becoming increasingly critical not just to developed countries but to developing ones as well \cite{yergin_2006}.
+Thus, assessing the energy security of different countries is important in order to understand the overall state of the international system.
 
-Energy security is defined by the International Energy Agency as ``the uninterrupted availability of energy sources at an affordable price''\citep{iea_2013}.
-As the world's energy demand continues to expand in the face of growing scarcity, energy security is becoming increasingly critical not just to developed countries but to developing ones as well \citep{yergin_2006}.
-Thus, assessing the energy security of different countries is important in order to understand the overall state of the international system.
-Despite this, there have been few published attempts to create formal models of energy security.
+Much of the current analysis of energy security focuses either on short-term vulnerabilities (in the time-scale of days and weeks \cite{Jewell_2011}), or on long-term investments in new technologies over the course of years and decades \cite{jacobson_2009}.
+The short-term analyses have largely concentrated on static indicators \cite{Jewell_2011}, with a particular focus on portfolio diversity (e.g.
+\cite{wu_2009,skea_2010,stirling_2010}), and do not incorporate the complex interactions that characterize the international system (\cite{cederman_1997,geller_2011}).
 
-One such model, MOSES (Model of Short-term Energy Security), was created within the International Energy Agency (IEA), the intergovernmental organization of the world's major oil importers created as a counterbalance to exporter cartel OPEC \citep{}.
-As the name indicates, MOSES focuses on short-term risks, defined as a timescale of days and weeks.
-Furthermore, MOSES assigns risk scores based on an impressive, but ultimately static, set of indicators.
-In particular, the indicator used to measure political stability is ``weighted average of political stability of suppliers''.
-However, the greater the diversity of suppliers (another input into the model), the greater the potential range of political stabilities.
-Furthermore, averaging ratings of political risk eliminates the complexity that characterises international political stabilitiy, ignoring the possibility of crisis and conflict contagion \citep{}.
-
-Other models of energy security focus on long-term timeframes, in years and decades. [[EXPAND MORE HERE]]
-
-In this paper, I present a novel model linking energy security with international political stability.
-
-\section{Model Description}
-
-\subsection{Overview}
-
-The model consists of countries, each of which has a set supply of and demand for crude oil.
-Countries are linked by a network of export-import relationships.
-Each month, some countries enter crisis with fixed probabilities; the exports and imports of countries in crisis are considered \emph{at risk}.
-If \textbf{contagion} is enabled, a crisis in one country may spread to its immediate geographic neighbors.
-If \textbf{assistance} is enabled, oil exporters may increase their production to balance a loss of secure oil by their major import partners.
-
-The model parameters currently remain fixed for each run: political instability, supply, demand, and trade relations are treated as constant.
-
-\subsection{Data Sources}
-The model inputs are empirical whenever possible.
-Data inputs are: oil trade relationships, country political instability, and country consumption of domestic oil production.
-Total supply and demand are derived from input data.
-
-The network of oil imports and exports comes from the United Nations-maintained COMTRADE database \citep{un_2013}.
-I extract all imports of crude petroleum (product code 2709) and other unprocessed petrolem (product code 2710) for 2012, the most recent year for which full data is available.
-Note that import data appears to be more reliable than export data: many major oil producers (particularly Saudi Arabia and Venezuela) only report export volume by region (e.g.
-Europe, North America); however, the majority of their trade partners report their imports from them.
-
-Political instability estimates are taken from the Economist Intelligence Unit's Political Index ratings \citep(eiu_2013).
-These are in turn estimated based on the methodology developed by the Political Instability Task Force \citep{goldstone_2005}, which could predict the onset of internal instability in a two-year period with 80\% accuracy.
-Thus, these scores are normalized such that a rating of 10 (the maximum) is associated with an 80\% probability of instability within a 24-month timeframe, translating into a [[FILLIN]] monthly probability of crisis onset.
-
-\section{Model Analysis}
-
-\section{Discussion}
-
-In this poster, I will present an initial implementation of a novel agent-based model linking crude oil energy security to inter- and intra-state crises and conflicts, through the mechanism of supply shocks \citep{kilian_2008}.
+In this poster, I will present an initial implementation of a novel agent-based model linking crude oil energy security to inter- and intra-state crises and conflicts, through the mechanism of supply shocks \cite{kilian_2008}.
 The model operates at the meso (months-years) timescale, and allows countries to react to changes in the oil supply.
 It can generate ranges of notional near-future trajectories of global and local energy security, providing risk estimates for different countries.
-The model attempts to capture the emergent consequences of conflict contagion \citep{black_2013}, temporary production increases and varying levels of stockpiles.
+The model attempts to capture the emergent consequences of conflict contagion \cite{black_2013}, temporary production increases and varying levels of stockpiles.
 It can also be used to explore different scenarios, from particular crises to realignments of major trade relationships.
@@
 The agents in the model are geospatially located countries, acting in monthly timesteps.
-Countries experience exogenous crises based on probabilities derived from empirical estimates (such as \citep{goldstone_2005,eiu_2013,ward_2013}).
-They are linked geographically, and by oil trade relationships obtained via the COMTRADE database \citep{un_2013}.
+Countries experience exogenous crises based on probabilities derived from empirical estimates (such as \cite{goldstone_2005,eiu_2013,ward_2013}).
+They are linked geographically, and by oil trade relationships obtained via the COMTRADE database \cite{un_2013}.
 As countries experience crises, their participation in the  international oil system is placed at risk, changing the volume of `secure' imports or exports of other countries and  potentially triggering supply or demand shocks.
@@
 
- \bibliographystyle{chicago}
-\bibliography{energysecurity}
+\bibliographystyle{splncs}
+\bibliography{sbp_bib}
 

### 2013-11-22 Writeups/CSS739 Paper/Paper.tex
Paper update

@@
 Data inputs are: oil trade relationships, country political instability, and country consumption of domestic oil production.
-Total supply and demand are derived from input data.
+Total supply and demand are derived largely from input data.
 
@@
 
-Political instability estimates are taken from the Economist Intelligence Unit's Political Index ratings \citep(eiu_2013).
+Political instability estimates are taken from the Economist Intelligence Unit's Political Index ratings \citep{eiu_2013}.
 These are in turn estimated based on the methodology developed by the Political Instability Task Force \citep{goldstone_2005}, which could predict the onset of internal instability in a two-year period with 80\% accuracy.
 Thus, these scores are normalized such that a rating of 10 (the maximum) is associated with an 80\% probability of instability within a 24-month timeframe, translating into a [[FILLIN]] monthly probability of crisis onset.
+
+For the majority of countries, I assume that supply and demand are equal to total export and imports, respectively.
+For the top ten oil producers, I use US Energy Information Administration (EIA) estimates \citep{} for those countries' consumption of their own domestic production.
+
+\subsection{Formal Description}
+
+The model consists of \textbf{Country} agents, linked by two networks: a directed \textbf{trade network}, and an undirected \textbf{geographic adjacency} network, where the country agents are the network nodes.
+
+Model variables may be \emph{data-derived}, \emph{parameters} that are fixed for each run and may be changed between them, or \emph{states} of the agents and edges that vary during the model.
+
+\subsubsection{Agent Description}
 

### 2013-11-24 Writeups/CSS739 Paper/Paper.tex
Parameterized
Made spare capacity and power-law exponent parameterized. Randomized
activation order.

@@
 	\usepackage{natbib}
+	\usepackage{mathtools}
+	\usepackage{amsmath}
+	\usepackage{algorithm2e}
 
@@
 The model consists of countries, each of which has a set supply of and demand for crude oil.
-Countries are linked by a network of export-import relationships.
+Countries are linked geographically, and by a network of export-import relationships.
 Each month, some countries enter crisis with fixed probabilities; the exports and imports of countries in crisis are considered \emph{at risk}.
@@
 
-The model consists of \textbf{Country} agents, linked by two networks: a directed \textbf{trade network}, and an undirected \textbf{geographic adjacency} network, where the country agents are the network nodes.
+The model consists of \textbf{Country} agents, posessing several characteristics and linked by two networks: a directed \textbf{trade network}, and an undirected \textbf{geographic adjacency} network, where the country agents are the network nodes.
 
-Model variables may be \emph{data-derived}, \emph{parameters} that are fixed for each run and may be changed between them, or \emph{states} of the agents and edges that vary during the model.
+Model variables may be \emph{fixed} based on data, \emph{parameters} that are fixed for each run and may be changed between them, or \emph{states} of the agents and edges that vary during the model.
 
-\subsubsection{Agent Description}
+\subsubsection{Networks}
 
-\section{Model Analysis}
+Agents are connected by two networks, which we can also express as matrices.
+Let $\mathbf{G}$ be the geographic adjacency matrix, such that $g_{i,j}=g_{j,i}=1$ when counties $i$ and $j$ are geographically contiguous or near-contiguous, and $0$ otherwise.
+
+Let $\mathbf{E}$ be the trade matrix, such that $e_{i,j}$ denotes the volume of oil exports from $i$ to $j$.
+Note that if the assistance submodel is turned on, the values of $\mathbf{E}$ may change over time.
+
+\subsubsection{Agent parameters and variables}
+
+Each agent represents an independent country.
+Agents are associated with the \textbf{geometry} of their position on a world map, which determines their neighbors.
+They are described by two submodels: an oil model, and a crisis model.
+
+\paragraph{Oil Submodel}
+\begin{itemize}
+	\item \textbf{Domestic Share} of the country's oil demand satisfied by domestic production.
+For most countries, this defaults to 0; for the top ten oil producers, it is set based on US Energy Information Administration data \citep{}.
+	\item \textbf{Total Imports} is the sum of the current volume of all oil imports. 
+	\item \textbf{Total Demand} is the sum of current imported and domestic oil consumed.
+Since the EIA reports the share of total consumption that is domestically-sourced, demand is calculated as: $$\text{Total Demand}_i = \sum_j e_{j,i} \cdot (1 - \text{Domestic Share})^{-1}$$
+	With $Total Imports$ set at the initial baseline. If current imports increase (as explained below) then supply will exceed demand. 
+	\item \textbf{Total Exports} is the sum of current exported oil to all other countries.
+\end{itemize}
+
+\paragraph{Crisis Submodel}
+\begin{itemize}
+	\item \textbf{Instability} is a measure of the country's political instability on a 0-10 scale, and determines the probability of a crisis or conflict -- a production shock.
+It is obtained from \citet{eiu_2013}.
+	\item \textbf{In Crisis?} is simply a boolean variable determining whether or not a country is currently experiencing a crisis or conflict.
+	\item \textbf{Crisis Length} is the number of months remaining in a crisis, if one is currently in process.
+When a country enters crisis, this number is drawn from a power law distribution, as described below.
+\end{itemize}
+
+\paragraph{Assistance Submodel}
+\begin{itemize}
+	\item \textbf{Increasing Production} is a boolean of whether or not the country is currently in a state of temporarily-increased production.
+	\item \textbf{Increasing Production For} stores the country being assisted. I assume that countries target their assistance at one other country.
+		\item \textbf{Crisis Length} is the number of months remaining in a crisis, if one is currently in process.
+When a country enters crisis, this number is drawn from a power law distribution, as described below.
+			\item \textbf{Total Capacity} is the country's capacity to rapidly increase oil for export above its inital baseline.
+Defaults to 1.1 (that is, countries can rapidly increase output by 10\%) based on consultations with subject-matter experts.
+\end{itemize}
+
+\subsubsection{Agent Behavior}
+
+Each tick of the model, agents are activated in random order, as follows:
+
+\begin{algorithm}[H]
+	\eIf{not in crisis} {
+		enter crisis with $Pr(crisis) \propto instability$ \;
+		\If{entered crisis}{
+			Crisis Length $\mathtt{\sim}$  Power Law \;
+			\If{Contagion turned on} {
+				\For{each neighboring country} {
+					Check whether neighbor enters crisis \;
+			}
+		}}}{
+		$Crisis Length = Crisis Length - 1$
+		}
+	\If{Assistance turned on} {
+		\eIf{Assisting another country} {
+			\If{Assisted country no longer in supply shock} {
+				End assistance\;
+			}
+		} {
+			\For{each out-neighbor} {
+				\If{Neighbor in supply shock} {
+					Assist neighbor with $Pr = \frac{Exports to neighbor}{Total exports}$
+				}
+			}
+		}
+	}	
+\end{algorithm}
+
+Once all agents have acted, their energy-security metrics are computed, as described below.
+
+\subsubsection{Energy Security Metrics}
+
+Let $\delta(i)$ be the \emph{Security indicator} function, defined as:
+\[
+\delta(i) = \begin{dcases*}
+	1 & when $i$ is not in crisis\\
+	0 & when $i$ is in crisis
+\end{dcases*}
+\]
+
+Country-level indicators are:
+
+\begin{itemize}
+	\item \textbf{Supply Ratio} is the fraction of the country's current demand being met by \emph{secure} imports -- that is, imports from countries that are not in crisis. Formally:
+	$$
+		\text{Supply Ratio}_i = \frac{\text{Total Demand}_i}{\sum_{j}e_{j,i}\delta(j)}
+	$$
+\item \textbf{Demand Ratio} is the fraction of the country's current exports going to countries that are not in crisis -- secure exports. 
+$$
+\text{Demand Ratio}_i = \frac{\sum_{j}e_{i,j}\delta(j)}{\sum_{j}e_{i,j}}
+$$
+\end{itemize}
+
+The model also computes similar global ratios:
+\begin{itemize}
+	\item \textbf{Global Supply Ratio} is the all country's current demand being met by secure imports:
+	$$
+		\text{Global Supply Ratio} = \frac{\sum_i\text{Total Demand}_i}{\sum_i\sum_{j}e_{j,i}\delta(j)}
+	$$
+	\item \textbf{Global Demand Ratio} is the fraction of worldwide exports going to countries that are not in crisis:
+	$$
+		\text{Global Demand Ratio} = \frac{\sum_i \text{Total Demand}_i\delta(i)}{\sum_i\sum_{j}e_{i,j}}
+	$$
+	\item \textbf{Global Overall Ratio} is the ratio between secure supply and secure demand:
+	$$
+		\text{Global Overall Ratio} = \frac{\sum_i \text{Total Demand}_i\delta(i)}{\sum_i\sum_{j}e_{j,i}\delta(j)}
+	$$
+\end{itemize}
+
+\section{Model Results and Analysis}
 

### 2013-11-24 Country.java

@@ -26,6 +26,7 @@ public class Country implements Steppable {
 	double domesticShare;
 	double totalDemand;
 	
+	double spareCapacity;
 	double totalCapacity;
 	
 	double supplyRatio;
@@ -121,7 +122,8 @@ public class Country implements Steppable {
 		// PLACEHOLDER ASSUMPTIONS:
 		// Oil consumption from imports only 
 		// Excess capacity 10% of current.
-		totalCapacity = 1.1 * totalExports;
+		spareCapacity = world.defaultExcessCapacity;
+		totalCapacity = (1 + spareCapacity) * totalExports;
 		
 	}
 	
@@ -169,7 +171,7 @@ public class Country implements Steppable {
 			inCrisis = true;
 			if (world.contagion) spreadContagion();
 			//crisisLength = Math.ceil(4 * Math.exp(world.random.nextGaussian()));
-			crisisLength = Math.ceil(world.util.randomPowerLaw(1, 800, -1.37));
+			crisisLength = Math.ceil(world.util.randomPowerLaw(1, 800, world.conflictExponent));
 		}
 	}
 	
@@ -258,6 +260,11 @@ public class Country implements Steppable {
 		domesticShare = newShare;
 		totalDemand = totalImports * 1.0/(1 - domesticShare);
 	}
+	public double getSpareCapacity() {return spareCapacity;}
+	public void setSpareCapacity(double newCapacity) {
+		spareCapacity = newCapacity;
+		totalCapacity = (1 + spareCapacity) * totalExports;
+	}
 	public double getTotalDemand() {return totalDemand;}
 	public boolean getInCrisis() {return inCrisis;}
 	public void setInCrisis(boolean crisis) {inCrisis = crisis;}
@@ -273,6 +280,9 @@ public class Country implements Steppable {
 	public void setInstability(double instability) {
 		this.instability = (instability/10.0) * 0.065;
 	}
+	
+	
+	
 	public XYSeries getRatioSeries() {return supplyRatioSeries;}
 
 }

### 2013-11-24 RiskWorld.java

@@ -5,11 +5,14 @@ import java.io.File;
 import java.io.FileReader;
 
 import java.net.URL;
+import java.util.ArrayList;
+import java.util.Collections;
 import java.util.HashMap;
 
 import com.vividsolutions.jts.geom.Envelope;
 
 import sim.engine.SimState;
+import sim.engine.Steppable;
 import sim.field.geo.GeomVectorField;
 import sim.field.network.Network;
 import sim.io.geo.ShapeFileImporter;
@@ -19,8 +22,8 @@ import sim.util.geo.MasonGeometry;
 @SuppressWarnings("serial")
 public class RiskWorld extends SimState {
 	// Model Parameters:
-	int worldHeight = 400;
-	int worldWidth = 900;
+	int worldHeight = 600;
+	int worldWidth = 1200;
 	// Parameters
 	// ==================================
 	double minEdgeToDisplay = 40000000; // Min export volume to show on the map.
@@ -28,6 +31,8 @@ public class RiskWorld extends SimState {
 	double shockThreshold = 2; // The demand/supply threshold required for a supply shock.
 	boolean contagion = false; // Flag to determine whether or not conflict contagion is active.
 	boolean assistance = true;
+	double conflictExponent = -1.37;
+	double defaultExcessCapacity = 0.1;
 	
 	// Storage
 	HashMap<String,Country> allCountries;
@@ -77,11 +82,21 @@ public class RiskWorld extends SimState {
 		//schedule.clear();
 		schedule.reset();
 		for (Country c : allCountries.values()) {
-			schedule.scheduleRepeating(c);
+			//schedule.scheduleRepeating(c);
 			c.initIndustry();
 			c.inCrisis = false;
 			c.crisisLength = -1;
 		}
+		schedule.scheduleRepeating(new Steppable() {
+			public void step(SimState state) {
+				ArrayList<String> names = new ArrayList<String>(allCountries.keySet());
+				Collections.shuffle(names);
+				for (String name : names) {
+					Country current = allCountries.get(name);
+					current.step(state);
+				}
+			}
+		});
 		tm = new TradeMonitor(this);
 		schedule.scheduleRepeating(tm);
 	}
@@ -241,7 +256,19 @@ public class RiskWorld extends SimState {
 	public void setContagion(boolean flag) {contagion = flag;}
 	public boolean getAssistance() { return assistance; }
 	public void setAssistance(boolean assistance) { this.assistance = assistance;}
[…]

### 2013-11-24 Utilities.java

@@ -4,6 +4,8 @@ package riskworld;
 import java.io.FileWriter;
 
 import java.util.ArrayList;
+import java.util.Arrays;
+import java.util.Collections;
 
 import com.vividsolutions.jts.geom.Coordinate;
 import com.vividsolutions.jts.geom.LineString;

### 2013-11-30 Writeups/CSS739 Paper/Paper.tex
Paper updates, begin adding parameterized spare capacity

@@
 
+For the majority of countries, I assume that supply and demand are equal to total export and imports, respectively.
+For the top ten oil producers, I use US Energy Information Administration (EIA) estimates \citep{} for those countries' consumption of their own domestic production.
+
 Political instability estimates are taken from the Economist Intelligence Unit's Political Index ratings \citep{eiu_2013}.
@@
 
-For the majority of countries, I assume that supply and demand are equal to total export and imports, respectively.
-For the top ten oil producers, I use US Energy Information Administration (EIA) estimates \citep{} for those countries' consumption of their own domestic production.
+Crisis duration is drawn independently from onset.
+\citet{cioffi_2004} and others have argued that internal and external conflict durations follow a power law distribution, which I use here.
+Specifically, in order to accomodate the limitations of the MASON random-number generator, I use a uniform approximation \citep{weisstein_2013}, defined as:
+$$
+\left((x_{1}^{\alpha} - x_{0}^{\alpha})y + x_{0}^{\alpha}\right)^{\frac{1}{\alpha}} \sim \text{Power Law}
+$$
+where 
+$$Y \sim U[0,1]$$
+$$ x_0  \equiv \text{minimum}$$
+$$ x_1  \equiv \text{maximum}$$
+$$ \alpha \equiv \text{Coefficient}$$
+
+I calibrate the parameters from two datasets: \citet{PRIO} for armed conflict, and \citep{hendrix_2013}, the Social Conflict in Africa Database, which includes both armed and lower-level social conflicts.
+I subset the PRIO dataset for conflicts in Africa and merge it with SCAD, eliminating duplicates and events below the severity of a general strike, in order to obtain as complete as possible a set of crises and conflicts that may be associated with an oil production  shock.
+I find the duration of each event, and fit a power law to the resulting distribution, resulting in an estimated coefficient of $\mathbf{-1.37}$.
 
@@
 \begin{algorithm}[H]
+	\caption{main loop}
 	\eIf{not in crisis} {
+			Call crisis submodel \;
+		}{
+			$Crisis Length = Crisis Length - 1$ \;
+			\If {Crisis Length == 0} { End crisis \; }
+		}
+	\If{Assistance turned on} {
+		Call assistance submodel \;
+	}	
+\end{algorithm}
+
+\begin{algorithm}[H]
+	\caption{crisis submodel}
 		enter crisis with $Pr(crisis) \propto instability$ \;
 		\If{entered crisis}{
-			Crisis Length $\mathtt{\sim}$  Power Law \;
+			Crisis Length $\mathtt{\sim}  \text{Power Law}$ \;
 			\If{Contagion turned on} {
 				\For{each neighboring country} {
-					Check whether neighbor enters crisis \;
+					Call neighbor's crisis submodel \;
+			}}}
+\end{algorithm}
+
+\begin{algorithm}[H]
+	\caption{assistance submodel}
+	\eIf{Currently assisting another country} {
+			\If{Assisted country no longer in supply shock} {
+				\tcp{Ending assistance}
+				\For{each export edge} {
+					edge volume = baseline edge volume;\
 				}
-		}}}{
-		$Crisis Length = Crisis Length - 1$
-		}
-	\If{Assistance turned on} {
-		\eIf{Assisting another country} {
-			\If{Assisted country no longer in supply shock} {
-				End assistance\;
 			}
@@
 				\If{Neighbor in supply shock} {
-					Assist neighbor with $Pr = \frac{Exports to neighbor}{Total exports}$
+					Assist neighbor with $Pr = \frac{\text{Exports to neighbor}}{\text{Total exports}}$
+				}
+			}
+			\If{Assisting neighbor} {
+				\tcp{Increasing output}
+				\For{each export edge} {
+					edge volume = edge volume * (1 + excess capacity)\;
 				}
@@
 		}
-	}	
+
 \end{algorithm}
@@
 
+\subsection{Validation}
+
+\subsection{Global Outcome Range}
+
+I run the model for 100 iterations for each permutation of the Contagion and Assistance flags, 400 iterations in total; each iteration was run for 60 ticks, or 5 simulated years.
+I collect the country-level and global indicators for each month of each iteration.
+
+Figure X shows a histogram of the Global Overall Ratio (the ratio between secure demand and supply) across all iteration-months.
+It appears to approximate a skew normal distribution, centered on 1 (a balance between supply and demand) but with a longer tail to the right -- supply shocks are more likely than demand shocks, and the most extreme supply shocks the model generates, though low-probability, are more extreme than any demand shocks.
+
+We can also examine the model iterations separately.
+Initially, we can characterize each iteration by the mean and variance of each indicator, as shown in Figure X.
+A high Overall or Supply mean ratio may indicate overall higher oil insecurity, but if the variance is low, it is `stably' insecure; in contrast, the mean may be close to 1 or even below it, but high variance would indicate severe uncertainty and fluctuation -- a distinct type of insecurity.
+Note that the overall correlation between mean and variance is XX, indicating that they are not directly linked to one another.
+
+We can further look at the trace of the indicators within each iteration, as they change from tick to tick.
+Figure X shows the traces of the 400 model runs overlaid on one another, and indicates some clear patterns.
+The majority of runs stay within a fairly narrow band, indicating similar levels of stability.
+Spikes outside the band, when they occur, tend to be brief.
+Note, however, that there are some longer-term spikes -- shocks which persist for multiple months, yielding values that become at least temporary baselines, even seeing fluctuations around them.
+
+\subsection{Contagion and Assistance}
+
+In order to disaggregate the effects of the Contagion and Assistance model parameters, we examine their results separately.
+Figure XX shows histograms of the scenario-months for Contagion and Assistance, respectively.
+The Contagion comparison indicates that conflict contagion does not appear to change the mode of the ratio distribution, which for both remains very close to 1.
+However, contagion does increase the variance of the distribution, making extreme outcomes more likely -- both to the right, indicating supply shocks, but also to the left, indicating demand shocks.
+
+Similarly, the modes of the distributions of ratios for scenario-months where Assistance is set to on and off are similarly both close to 1.
+
 \section{Discussion}

### 2013-11-30 Country.java

@@ -52,7 +52,9 @@ public class Country implements Steppable {
 		this.world = world;
 		this.name = name;
 		// Instability=100 <=> Approx. 80% chance of crisis occurring once in 24 tests.
-		this.instability = (instability/10.0) * 0.065; 
+		this.instability = (instability/10.0) * 0.065;
+		
+		spareCapacity = world.defaultExcessCapacity;
 		crisisCheckCounter = 0;
 	}
 	
@@ -120,9 +122,7 @@ public class Country implements Steppable {
 		
 		// TODO: Update with real data
 		// PLACEHOLDER ASSUMPTIONS:
-		// Oil consumption from imports only 
 		// Excess capacity 10% of current.
-		spareCapacity = world.defaultExcessCapacity;
 		totalCapacity = (1 + spareCapacity) * totalExports;
 		
 	}

### 2013-12-05 Writeups/CSS739 Paper/Paper.tex
Minor tweaks, paper work

@@
 
-One such model, MOSES (Model of Short-term Energy Security), was created within the International Energy Agency (IEA), the intergovernmental organization of the world's major oil importers created as a counterbalance to exporter cartel OPEC \citep{}.
-As the name indicates, MOSES focuses on short-term risks, defined as a timescale of days and weeks.
+Much of the analysis has focused on a portfolio-based approach.
+\citet{wu_2009} uses a portfolio diversity approach to analyze China's mix of crude oil and petrolem products.
+While the paper notes that ``Oil import-specific risk derives from events, such as political events, natural disaster and so on, that are not related to general movements in the international market'', but does not analyze the risks of such events directly.
+It does, however, address the relatively less significant risk of piracy along several shipping routes.
+\citet{stirling_2010} and \citet{skea_2010} take a broader, more theoretical approach, and develops a multicriteria methodology for evaluating the diversity of an entire energy portfolio as it contributes to energy security.
+However, changing an entire energy portfolio takes a great deal of time -- it is a long-term process, measured in years or decades.
+Similarly, \citet{jacobson_2009} evaluates a set of possible future courses of action to jointly address climate, environmental and energy-security concerns -- again, courses of actions only implementable in the long term.
+
+Short-term energy security, in contrast, appears to have been little-addressed in the literature.
+However, an important short-term analysis, MOSES (Model of Short-term Energy Security), was created within the International Energy Agency (IEA), the intergovernmental organization of the world's major oil importers created as a counterbalance to exporter cartel OPEC \citep{}.
+As the name indicates, MOSES focuses on short-term risks, which it defines as a timescale of days and weeks.
 Furthermore, MOSES assigns risk scores based on an impressive, but ultimately static, set of indicators.
@@
 
-Other models of energy security focus on long-term timeframes, in years and decades. [[EXPAND MORE HERE]]
+In fact, as \citet{hamilton_2003} describes, oil supply shocks have been driven caused by conflict involving oil producers, both domestic and international.
+\citet{hamilton_2003} argues that supply shock-driven changes to oil prices have a nonlinear effect on GDPs, while \citet{killian_2008} argues that the macroeconomic effects of the oil shocks themselves are small.
+Nevertheless, the price of oil is only an indirect indicator of energy security; if a substantial volume of oil cannot be extracted or exported, and hence is unavailable for consumption, it poses an energy security risk regardless of what the price of oil is doing.
 
-In this paper, I present a novel model linking energy security with international political stability.
+In this paper, I present a novel model linking energy security with political conflict.
+At the core of the model is the supply and demand for crude oil at the country level, both imported and domestically-produced.
+Supply and demand are placed at risk when countries experience crises and conflicts.
+The model operates on the meso timescale, of months and years rather than days or decades.
+It produces local and global time-series of oil security indicators with each iteration.
+By running multiple iterations with different parameters, I generate notional near-future scenarios.
+
+The model straddles the boundary between a complicated cellular automata and a simple agent-based model.
+As we shall see, at its simplest configuration, the model is essentially stochastically sampling from the space of possible international crisis configurations and measuring their impact, with no agent decisionmaking or interaction.
+By incrementally adding more  elements (in particular, the possibility of conflict contagion and rapid increases of production to counterbalance supply shocks), we can estimate the impact of the additional complexity.
+
+In the remainder of the paper, I describe the model in more detail, its data sources, and its various submodels. I use the model to conduct several experiments, and describe the output.
 
@@
 The model consists of countries, each of which has a set supply of and demand for crude oil.
-Countries are linked geographically, and by a network of export-import relationships.
-Each month, some countries enter crisis with fixed probabilities; the exports and imports of countries in crisis are considered \emph{at risk}.
+Countries are linked geographically, and by a network of export-import relationships, and each country has a fixed supply of, and demand for, crude oil.
+Each month, some countries enter crisis at random; the exports and imports of countries in crisis are considered \emph{at risk}.
 If \textbf{contagion} is enabled, a crisis in one country may spread to its immediate geographic neighbors.
@@
 
-\subsection{Global Outcome Range}
+\subsection{Global Outcomes}
+
+\subsubsection{Outcome Range}
 
@@
 
-\subsection{Contagion and Assistance}
+\subsubsection{Contagion and Assistance}
 
@@
 Figure XX shows histograms of the scenario-months for Contagion and Assistance, respectively.
-The Contagion comparison indicates that conflict contagion does not appear to change the mode of the ratio distribution, which for both remains very close to 1.
-However, contagion does increase the variance of the distribution, making extreme outcomes more likely -- both to the right, indicating supply shocks, but also to the left, indicating demand shocks.
 
-Similarly, the modes of the distributions of ratios for scenario-months where Assistance is set to on and off are similarly both close to 1.
+Upon visual examination, the distributions with Contagion set to 'On' and 'Off' appear extremely similar.
+I conduct two statistical tests to determine whether they are in fact drawn from the same underlying distribution -- that is, whether there is no actual difference between them.
+Both the Kolmogorov-Smirnov  and Mann-Whitney tests indicate that the distributions are in fact different.
+Nevertheless, the differences are small.
+The mean ratio for the No Contagion scenario is 1.046, compared to 1.068 for the Contagion scenario.
+Both a parametric T-Test and nonparametric Wilcoxon Rank-Sum confirm that these means are indeed different, though the difference is small.
+Similarly, the variance for the No Contagion scenario is 0.045, while the Contagion variance is 0.051.
+Thus, conflict contagion pushes the scenarios away from the mean, yielding a wider range of outcomes.
+
+I repeat the same analysis for the the Assistance disabled and enabled scenarios.
+Again, while the distributions appear extremely similar, both the Kolmogorov-Smirnov and Mann-Whitney tests show that they are different.
+The mean and variance for Assistance Off are 1.062 and 0.05, respectively, compared to 1.053 and 0.046 when Assistance is enabled.
+This confirms with our intuition -- international assistance reduces the uncertainty of energy availability, and increases overall energy security, albeit by a small number.
+
+\subsection{Country-level analysis}
+
+In addition to the aggregate global indicators, the model also tracks the Supply and Demand ratios for each country.
+We can analyze this output in order to understand how different countries' risks differ from one another, and to identify countries at high or low risk.
+
+Figure XX shows a map of the overall risk of a major supply crisis, with the supply ratio rising above 2.
+Several things immediately jump out: several countries at the highest risk countries are major net oil exporters, including Canada, Mexico and Iran.
+While these countries derive a large fraction of their consumption from domestic production, they also rely on imports from abroad
+
+\subsection{Sensitivity Analysis}
 
 \section{Discussion}
-
-In this poster, I will present an initial implementation of a novel agent-based model linking crude oil energy security to inter- and intra-state crises and conflicts, through the mechanism of supply shocks \citep{kilian_2008}.
-The model operates at the meso (months-years) timescale, and allows countries to react to changes in the oil supply.
-It can generate ranges of notional near-future trajectories of global and local energy security, providing risk estimates for different countries.
-The model attempts to capture the emergent consequences of conflict contagion \citep{black_2013}, temporary production increases and varying levels of stockpiles.
-It can also be used to explore different scenarios, from particular crises to realignments of major trade relationships.
-
-The agents in the model are geospatially located countries, acting in monthly timesteps.
-Countries experience exogenous crises based on probabilities derived from empirical estimates (such as \citep{goldstone_2005,eiu_2013,ward_2013}).
-They are linked geographically, and by oil trade relationships obtained via the COMTRADE database \citep{un_2013}.
-As countries experience crises, their participation in the  international oil system is placed at risk, changing the volume of `secure' imports or exports of other countries and  potentially triggering supply or demand shocks.
-
-Countries respond to the changing energy security environment in several ways.
-Exporters may choose to temporarily increase production in response to supply shocks elsewhere, while countries that are members of the International Energy Agency may vote to release their emergency stockpiles.
-The model's behavior is verified against previous qualitative and quantitative analysis, and in consultation with subject-matter experts.
-
-The poster will describe the model's methodology and input data, and provide a visual representation of the model's estimated energy security risk for every country.
-It will provide a visual example of a representative model run, and illustrate the results of a sweep of the model's behavior and parameter space.
-Finally, it will describe future work, incorporating different energy sources, shipping chokepoints, and dynamic network updating.
 

### 2013-12-05 Country.java

@@ -54,7 +54,7 @@ public class Country implements Steppable {
 		// Instability=100 <=> Approx. 80% chance of crisis occurring once in 24 tests.
 		this.instability = (instability/10.0) * 0.065;
 		
-		spareCapacity = world.defaultExcessCapacity;
+		spareCapacity = world.defaultSpareCapacity;
 		crisisCheckCounter = 0;
 	}
 	
@@ -274,8 +274,8 @@ public class Country implements Steppable {
 	public double getLocalRatio() {return supplyRatio;}
 	
 	public double getInstability() {
-		return instability;
-		//return instability/0.065 * 10;
+		//return instability;
+		return instability/0.065 * 10;
 	}
 	public void setInstability(double instability) {
 		this.instability = (instability/10.0) * 0.065;

### 2013-12-05 RiskWorld.java

@@ -32,7 +32,7 @@ public class RiskWorld extends SimState {
 	boolean contagion = false; // Flag to determine whether or not conflict contagion is active.
 	boolean assistance = true;
 	double conflictExponent = -1.37;
-	double defaultExcessCapacity = 0.1;
+	double defaultSpareCapacity = 0.1;
 	
 	// Storage
 	HashMap<String,Country> allCountries;
@@ -113,10 +113,14 @@ public class RiskWorld extends SimState {
 				String name = row[0];
 				double instability = Double.parseDouble(row[3]);
 				double domesticShare = 0;
-				if (row.length == 5) 
+				double spareCapacity = this.defaultSpareCapacity;
+				if (row.length >= 5) 
 					domesticShare = Double.parseDouble(row[4]);
+				if (row.length >= 6)
+					spareCapacity = Double.parseDouble(row[5]);
 				Country newCountry = new Country(this, name, instability);	
 				newCountry.domesticShare = domesticShare;
+				newCountry.spareCapacity = spareCapacity;
 				allCountries.put(name, newCountry);
 				schedule.scheduleRepeating(newCountry);
 			}
@@ -258,14 +262,14 @@ public class RiskWorld extends SimState {
 	public void setAssistance(boolean assistance) { this.assistance = assistance;}
 	public double getConflictExponent() {return conflictExponent;}
 	public void setConflictExponent(double exponent) {conflictExponent = exponent;}
-	public double getDefaultExcessCapacity() {return defaultExcessCapacity;}
+	public double getDefaultExcessCapacity() {return defaultSpareCapacity;}
 	public void setDefaultExcessCapacity(double capacity) {
-		double oldCapacity = defaultExcessCapacity;
+		double oldCapacity = defaultSpareCapacity;
 		for (Country c : allCountries.values()) {
 			if (oldCapacity == c.spareCapacity)
 				c.setSpareCapacity(capacity);
 		}
-		defaultExcessCapacity = capacity;
+		defaultSpareCapacity = capacity;
 	}
 	public double[] getAllRatios() {return allSupplyRatios;}
 	

### 2013-12-05 SensitivityAnalysis.java

@@ -8,39 +8,38 @@ import org.jfree.data.xy.XYSeries;
 import com.google.gson.Gson;
 
 /**
- * Class for running the main model experiment in bulk mode, saving the outputs.
- * @author dmasad
+ * Class for running the sensitivity analysis, saving global output and parameters.
  *
  */
-public class BulkRunner {
+public class SensitivityAnalysis {
 	
 	/**
-	 * Class to store the output of a single model run.
-	 *
+	 * Class to store the output of a single model run, for global variables only.
+	 * NOTE: Different from the Model Output defined in BulkRunner.
+	 * 
 	 */
 	static class ModelOutput {
 		boolean contagion;
 		boolean assistance;
+		double conflictExponent;
+		double defaultSpareCapacity;
+		double shockThreshold;
+		
 		double globalSupplyRatio[];
 		double globalDemandRatio[];
 		double globalOverallRatio[];
-		HashMap<String, double[]> supplyRatios;
-		HashMap<String, double[]> demandRatios;
 		
 		ModelOutput(RiskWorld world) {
 			contagion = world.contagion;
 			assistance = world.assistance;
+			conflictExponent = world.conflictExponent;
+			defaultSpareCapacity = world.defaultSpareCapacity;
+			shockThreshold = world.shockThreshold;
 			TradeMonitor tm = world.tm;
 			int ticks = (int)world.schedule.getSteps();
 			globalSupplyRatio = convertXYSeries(tm.globalSupplyRatio);
 			globalDemandRatio = convertXYSeries(tm.globalDemandRatio);
 			globalOverallRatio = convertXYSeries(tm.globalOverallRatio);
-			supplyRatios = new HashMap<String, double[]>();
-			demandRatios = new HashMap<String, double[]>();
-			for (Country c : world.allCountries.values()) {
-				supplyRatios.put(c.name, convertXYSeries(c.supplyRatioSeries));
-				demandRatios.put(c.name, convertXYSeries(c.demandRatioSeries));
-			}
 			
 		}
 		
@@ -80,32 +79,32 @@ public class BulkRunner {
 	 * 
 	 */
 	public static void main(String[] args) {
-		int n = 250; // Number of model runs:
+		int n = 10; // Number of model runs per permutation
 		int j = 0; // Current iteration counter
 		int numSteps = 60;
 		Gson gson = new Gson();
[…]

### 2013-12-07 Country.java
Fixed major error in supplyRatio calculation

@@ -137,7 +137,7 @@ public class Country implements Steppable {
 			Country neighbor = (Country)e.getFrom();
 			if (!neighbor.inCrisis) currentImports += ((TradeEdge)e.getInfo()).currentSize;
 		}
-		supplyRatio = totalDemand/currentImports;
+		supplyRatio = totalDemand/(currentImports + (domesticShare*totalDemand));
 		if (supplyRatio > world.shockThreshold) {
 			supplyShock = true;
 			if (world.verbose) System.out.println(name + " experiencing supply shock");
@@ -271,7 +271,7 @@ public class Country implements Steppable {
 	public double getCrisisLength() {return crisisLength;}
 	public void setCrisisLength(double length) {crisisLength = length;}
 	public int getCrisisCheckCounter() {return crisisCheckCounter;}
-	public double getLocalRatio() {return supplyRatio;}
+	public double getSupplyRatio() {return supplyRatio;}
 	
 	public double getInstability() {
 		//return instability;

### 2013-12-07 TradeMonitor.java
Updated Trade Monitor to fix TotalDemand bug

@@ -30,8 +30,8 @@ public class TradeMonitor implements Steppable {
 	public TradeMonitor(RiskWorld world) {
 		this.world = world;
 		for (Country c : world.allCountries.values()) {
-			totalSupply += c.totalExports;
-			totalDemand += c.totalImports;
+			totalSupply += (c.totalExports + c.domesticShare*c.totalDemand);
+			totalDemand += c.totalDemand;
 		}
 	}
 	
@@ -41,6 +41,8 @@ public class TradeMonitor implements Steppable {
 		allSupplyRatios = new double[world.allCountries.size()];
 		int i = 0;
 		for (Country c : world.allCountries.values()) {
+			currentDemand += (c.domesticShare*c.totalDemand);
+			currentSupply += (c.domesticShare * c.totalDemand);
 			if (!c.inCrisis) {
 				currentSupply += c.totalExports;
 				currentDemand += c.totalImports;

### 2013-12-08 Writeups/CSS739 Paper/Paper.tex
Paper updates

@@
 
-I run the model for 100 iterations for each permutation of the Contagion and Assistance flags, 400 iterations in total; each iteration was run for 60 ticks, or 5 simulated years.
+I run the model for 250 iterations for each permutation of the Contagion and Assistance flags, 1,000 iterations in total; each iteration was run for 60 ticks, or 5 simulated years.
 I collect the country-level and global indicators for each month of each iteration.
@@
 Figure X shows a histogram of the Global Overall Ratio (the ratio between secure demand and supply) across all iteration-months.
-It appears to approximate a skew normal distribution, centered on 1 (a balance between supply and demand) but with a longer tail to the right -- supply shocks are more likely than demand shocks, and the most extreme supply shocks the model generates, though low-probability, are more extreme than any demand shocks.
+It appears to approximate a skew normal distribution, centered very close to 1 (a balance between supply and demand) but with a longer tail to the right -- supply shocks are more likely than demand shocks, and the most extreme supply shocks the model generates, though low-probability, are more extreme than any demand shocks.
 
@@
 Upon visual examination, the distributions with Contagion set to 'On' and 'Off' appear extremely similar.
-I conduct two statistical tests to determine whether they are in fact drawn from the same underlying distribution -- that is, whether there is no actual difference between them.
+I conduct several statistical tests to determine whether they are in fact drawn from the same underlying distribution -- that is, whether there is no actual difference between them.
 Both the Kolmogorov-Smirnov  and Mann-Whitney tests indicate that the distributions are in fact different.
 Nevertheless, the differences are small.
-The mean ratio for the No Contagion scenario is 1.046, compared to 1.068 for the Contagion scenario.
-Both a parametric T-Test and nonparametric Wilcoxon Rank-Sum confirm that these means are indeed different, though the difference is small.
-Similarly, the variance for the No Contagion scenario is 0.045, while the Contagion variance is 0.051.
+The mean ratio for the No Contagion scenario is 1.037, compared to 1.049 for the Contagion scenario.
+Both a parametric T-Test and nonparametric Wilcoxon Rank-Sum confirm that these means are indeed different, and neither the traditional (frequentist) not Bayesian confidence intervals overlap.
+Similarly, the variance for the No Contagion scenario is 0.025, while the Contagion variance is 0.03.
 Thus, conflict contagion pushes the scenarios away from the mean, yielding a wider range of outcomes.
@@
 I repeat the same analysis for the the Assistance disabled and enabled scenarios.
-Again, while the distributions appear extremely similar, both the Kolmogorov-Smirnov and Mann-Whitney tests show that they are different.
-The mean and variance for Assistance Off are 1.062 and 0.05, respectively, compared to 1.053 and 0.046 when Assistance is enabled.
-This confirms with our intuition -- international assistance reduces the uncertainty of energy availability, and increases overall energy security, albeit by a small number.
+However, in this case, the statistical tests cannot reject the hypothesis that the resulting distributions are the same, and the means are within each other's confidence intervals.
 
@@
 
-Figure XX shows a map of the overall risk of a major supply crisis, with the supply ratio rising above 2.
-Several things immediately jump out: several countries at the highest risk countries are major net oil exporters, including Canada, Mexico and Iran.
-While these countries derive a large fraction of their consumption from domestic production, they also rely on imports from abroad
+Figure XX shows a map of the overall risk of a major supply crisis, with the supply ratio rising above 1.5 (that is, demand exceeds secure supply by at least 50\%).
+Several things immediately jump out: major oil producers appear to be at the least risk, as their consumption of domestic production reduces their exposure to import disruption.
+Russia stands out as the one major producer to face a significant supply risk.
+The countries of the so-called `Global South' appears to face higher risks than the more developed world, likely indicating a lower diversity of import sources.
+
+Figure XX shows a similar map for demand risk, and indicates the likelihood of total possible exports exceed secure exports by at least 50\%.
+While there are fewer countires with very low risk, there are fewer with extremely high risk as well.
+Interestingly, the major Middle East producers appear to face relatively low demand risk, likely indicating a robust set of trade relationships with realively stable countries.
+In contrast, the major non-US producers in the Americas -- Canada, Mexico and Venezuela -- appear to face higher demand risk.
+This appears to be due to the fact that their exports are disproportionately to the United States, rendering them vulnerable to US crises.
 
 \subsection{Sensitivity Analysis}
+
+As discussed above, the model accepts several input parameters, most of which are calibrated from data.
+Given that this calibration is only approximate, it is important to conduct a sensitivity analysis in order to identify whether the results are extremely sensitive to the input parameters -- the more sensitive they are, the more the model will be biased by missing or incomplete data, and the less reliable we expect it to be.
+The space of parameters and outputs to search is extremely large; for the purpose of this analysis, I will limit myself to two parameters: the exponential coefficient on the conflict length power law (calibrated from data, as described above) and the default spare production capacity, estimated from SME input.
+The output I examine is the Overall Ratio, as it provides a global summary of the model condition.
+
+In order to conduct the sensitivity analysis, I iterate over conflict coefficient values between -1.05 and -2.0, and default spare production capacities between 0 and 0.5, both in steps of 0.05.
+For each combination, I run the model 40 times, 10 iterations each for each combination of Contagion and Assistance, for a total of 2,200 model runs.
 

### 2013-12-08 Writeups/CSS739 Paper/Paper.tex
paper update

@@
 Much of the analysis has focused on a portfolio-based approach.
-\citet{wu_2009} uses a portfolio diversity approach to analyze China's mix of crude oil and petrolem products.
-While the paper notes that ``Oil import-specific risk derives from events, such as political events, natural disaster and so on, that are not related to general movements in the international market'', but does not analyze the risks of such events directly.
+\citet{wu_2009} uses portfolio diversity to analyze China's mix of crude oil and petrolem products.
+The paper notes that ``Oil import-specific risk derives from events, such as political events, natural disaster and so on, that are not related to general movements in the international market'', but does not analyze the risks of such events directly.
 It does, however, address the relatively less significant risk of piracy along several shipping routes.
@@
 Short-term energy security, in contrast, appears to have been little-addressed in the literature.
-However, an important short-term analysis, MOSES (Model of Short-term Energy Security), was created within the International Energy Agency (IEA), the intergovernmental organization of the world's major oil importers created as a counterbalance to exporter cartel OPEC \citep{}.
-As the name indicates, MOSES focuses on short-term risks, which it defines as a timescale of days and weeks.
+However, an important short-term analysis, MOSES (Model of Short-term Energy Security), was created within the International Energy Agency (IEA), the intergovernmental organization of the world's major oil importers established in part to protect their energy security.
+As the name of the model indicates, MOSES focuses on short-term risks, which it defines as a timescale of days and weeks.
 Furthermore, MOSES assigns risk scores based on an impressive, but ultimately static, set of indicators.
@@
 
-\subsection{Validation}
-
 \subsection{Global Outcomes}
@@
 We can further look at the trace of the indicators within each iteration, as they change from tick to tick.
-Figure X shows the traces of the 400 model runs overlaid on one another, and indicates some clear patterns.
+Figure X shows the traces of the 1,000 model runs overlaid on one another, and indicates some clear patterns.
 The majority of runs stay within a fairly narrow band, indicating similar levels of stability.
@@
 In order to disaggregate the effects of the Contagion and Assistance model parameters, we examine their results separately.
-Figure XX shows histograms of the scenario-months for Contagion and Assistance, respectively.
+Figure XX shows histograms of the Overall Ratio for the scenario-months for Contagion and Assistance, respectively.
 
-Upon visual examination, the distributions with Contagion set to 'On' and 'Off' appear extremely similar.
+Upon visual examination, the distributions with Contagion set to `On' and `Off' appear extremely similar.
 I conduct several statistical tests to determine whether they are in fact drawn from the same underlying distribution -- that is, whether there is no actual difference between them.
@@
 However, in this case, the statistical tests cannot reject the hypothesis that the resulting distributions are the same, and the means are within each other's confidence intervals.
+
+I conduct similar analyses for the other global ratios, for supply and demand, as well, which result in the same conclusion.
+The effects of Contagion are small but statistically significant, while Assistance has a far smaller and less significant effect.
 
@@
 The space of parameters and outputs to search is extremely large; for the purpose of this analysis, I will limit myself to two parameters: the exponential coefficient on the conflict length power law (calibrated from data, as described above) and the default spare production capacity, estimated from SME input.
-The output I examine is the Overall Ratio, as it provides a global summary of the model condition.
+I will only examine the global ratios, as proxies for the behavior of the overall model.
 
@@
 
+Figure XX shows the mean and variance of the Global Ratios produced for each iteration-tick at each combination of the parameters.
+Visual examination shows that there is not significant change in the Overall Ratio; the range of outputs is relatively small, and no combination of inputs produces significant outliers.
+The one input-output combination detectable is between the conflict coefficient and the variance.
+As the conflict coefficient increases (reducing the probably of long conflicts), the variance is reduced as well.
+However, for the Supply and Demand Ratios, there is a clear change driven by conflict coefficient: as the probability of long conflicts increases, the mean of the Supply Ratio goes up and of the Demand Ratio goes down.
+Note, however, that that there does not appear to be a clear critical point; the change relatively slow and linear.
+Note also that the calibrated value of the parameter is in the higher-variance, higher-risk zone.
+In other words, the model may be forecasting risk on the higher end.
+
 \section{Discussion}
+
+In this paper, I have described an agent-based model to study near-future international energy security, with a focus on crude oil imports and exports.
+The model is data-driven, ties energy security to political crises and conflicts, and can study the effects of conflict contagion and of short-term oil production increases.
+We can draw two sets of conclusions from the model and its outputs: on energy security itself, as well as on the challenges of modeling it, and modeling international systems more broadly.
+
+In terms of energy security, the model does not predict extreme shocks to be likely.
+Nevertheless, it shows that such extreme shocks may occur within the current configuration of the international oil system, and provides an estimate for their probability.
+The model provides evidence that consumption of domestic oil production provides a significant improvement in energy security, even in the absence of total energy independence.
+It also highlights the importance of diverse import and export partners.
+
+The model identifies the countries facing the greatest risks to their energy security, as well as making patterns in the risk visible -- particularly the high risk in the countries of the `Global South'.
+
+Analysis of the Contagion and Assistance parameters yields additional interesting insight.
+The effect of contagion is statistically significant but small -- perhaps suggesting the difficulty facing theongoing debate over the phenomenon in political science, but also that the debate's conclusions may not be very important to policy.
+More counterintuitively, perhaps, the Assistance parameter does not have a significant, measurable effect.
+Here is somewhere the model can generate insights that are not otherwise obvious -- specifically, that even if major oil producers can rapidly increase their output in response to crises, it is not enough to counterbalance the effects of those crises.
+
+From a technical perspective, the model presented here is extremely simple.
+The interaction between agents is minimal and largely indirect, and with both Contagion and Assistance turned off, the model is essentially stochastically sampling from the space of possible crises and durations.
+Nevertheless, it is sufficient to capture the core aspects of the international oil system and produces useful results.
+
+However, even such a simple model had extensive input requirements, with data coming from various sources of varying completeness and reliability.
+Expanding the model (for example, by adding additional energy sources, or incorporating oil refineries and processed petroleum) would significantly increase the data requirements.
+Adding dynamics and behavior pose an even greater challenge.
 

### 2013-12-11 Writeups/CSS739 Paper/Paper.tex
Added graphics

@@
 Thus, assessing the energy security of different countries is important in order to understand the overall state of the international system.
-Despite this, there have been few published attempts to create formal models of energy security.
+Despite this, there have been few attempts to create formal models of energy security.
 
-Much of the analysis has focused on a portfolio-based approach.
-\citet{wu_2009} uses portfolio diversity to analyze China's mix of crude oil and petrolem products.
+Much of the analysis to date has taken a portfolio-based approach.
+\citet{wu_2009} uses portfolio diversity to analyze China's mix of crude oil and processed petrolem products.
 The paper notes that ``Oil import-specific risk derives from events, such as political events, natural disaster and so on, that are not related to general movements in the international market'', but does not analyze the risks of such events directly.
@@
 \citet{stirling_2010} and \citet{skea_2010} take a broader, more theoretical approach, and develops a multicriteria methodology for evaluating the diversity of an entire energy portfolio as it contributes to energy security.
-However, changing an entire energy portfolio takes a great deal of time -- it is a long-term process, measured in years or decades.
-Similarly, \citet{jacobson_2009} evaluates a set of possible future courses of action to jointly address climate, environmental and energy-security concerns -- again, courses of actions only implementable in the long term.
+However, changing the composition of an entire energy portfolio takes a great deal of time -- it is a long-term process, measured in years or decades.
+Similarly, \citet{jacobson_2009} evaluates a set of possible future courses of action to jointly address climate, environmental and energy-security concerns -- again, courses of actions only implementable over the long term.
 
@@
 As the name of the model indicates, MOSES focuses on short-term risks, which it defines as a timescale of days and weeks.
-Furthermore, MOSES assigns risk scores based on an impressive, but ultimately static, set of indicators.
-In particular, the indicator used to measure political stability is ``weighted average of political stability of suppliers''.
+Furthermore, MOSES assigns risk scores based on a broad but ultimately static set of indicators.
+In particular, the indicator used to measure political stability is the ``weighted average of political stability of suppliers''.
 However, the greater the diversity of suppliers (another input into the model), the greater the potential range of political stabilities.
-Furthermore, averaging ratings of political risk eliminates the complexity that characterises international political stabilitiy, ignoring the possibility of crisis and conflict contagion \citep{}.
+Averaging ratings of political risk eliminates the complexity that characterises international political stabilitiy, ignoring the possibility of crisis and conflict contagion.
 
@@
 It produces local and global time-series of oil security indicators with each iteration.
-By running multiple iterations with different parameters, I generate notional near-future scenarios.
+By running multiple iterations with different parameters, I generate notional near-future scenarios of changes in international energy security.
 
@@
 
-In the remainder of the paper, I describe the model in more detail, its data sources, and its various submodels. I use the model to conduct several experiments, and describe the output.
+In the remainder of the paper, I describe the model in more detail, its data sources, and its various submodels.
+I use the model to conduct several experiments, describe the output, and conduct a sensitivity analysis.
+Finally, I discuss my conclucions, both regarding international energy security and the challenges and opportunities of this type of modeling.
 
@@
 The model consists of countries, each of which has a set supply of and demand for crude oil.
-Countries are linked geographically, and by a network of export-import relationships, and each country has a fixed supply of, and demand for, crude oil.
+Countries are linked geographically, and by a network of export-import relationships.
 Each month, some countries enter crisis at random; the exports and imports of countries in crisis are considered \emph{at risk}.
-If \textbf{contagion} is enabled, a crisis in one country may spread to its immediate geographic neighbors.
-If \textbf{assistance} is enabled, oil exporters may increase their production to balance a loss of secure oil by their major import partners.
+If \textbf{Contagion} is enabled, a crisis in one country may spread to its immediate geographic neighbors.
+If \textbf{Assistance} is enabled, oil exporters may increase their production to balance a loss of secure oil by their major import partners.
 
@@
 For the majority of countries, I assume that supply and demand are equal to total export and imports, respectively.
-For the top ten oil producers, I use US Energy Information Administration (EIA) estimates \citep{} for those countries' consumption of their own domestic production.
+For the top ten oil producers, I use US Energy Information Administration (EIA) estimates \citep{eia_2013} for those countries' consumption of their own domestic production.
 
@@
 \citet{cioffi_2004} and others have argued that internal and external conflict durations follow a power law distribution, which I use here.
-Specifically, in order to accomodate the limitations of the MASON random-number generator, I use a uniform approximation \citep{weisstein_2013}, defined as:
-$$
-\left((x_{1}^{\alpha} - x_{0}^{\alpha})y + x_{0}^{\alpha}\right)^{\frac{1}{\alpha}} \sim \text{Power Law}
-$$
-where 
-$$Y \sim U[0,1]$$
-$$ x_0  \equiv \text{minimum}$$
-$$ x_1  \equiv \text{maximum}$$
-$$ \alpha \equiv \text{Coefficient}$$
-
-I calibrate the parameters from two datasets: \citet{PRIO} for armed conflict, and \citep{hendrix_2013}, the Social Conflict in Africa Database, which includes both armed and lower-level social conflicts.
+I calibrate the parameters from two datasets: the UCDP/PRIO Armed Conflict Dataset \citep{lotta_2013} for military conflicts, and the Social Conflict in Africa Database \citep{hendrix_2013}, which includes both armed and lower-level social conflicts.
 I subset the PRIO dataset for conflicts in Africa and merge it with SCAD, eliminating duplicates and events below the severity of a general strike, in order to obtain as complete as possible a set of crises and conflicts that may be associated with an oil production  shock.
@@
 	\item \textbf{Domestic Share} of the country's oil demand satisfied by domestic production.
-For most countries, this defaults to 0; for the top ten oil producers, it is set based on US Energy Information Administration data \citep{}.
+For most countries, this defaults to 0; for the top ten oil producers, it is set based on US Energy Information Administration data \citep{eia_2013}.
 	\item \textbf{Total Imports} is the sum of the current volume of all oil imports. 
@@
 	\item \textbf{Crisis Length} is the number of months remaining in a crisis, if one is currently in process.
-When a country enters crisis, this number is drawn from a power law distribution, as described below.
+When a country enters crisis, this number is drawn from a power law distribution.
+Specifically, in order to accomodate the limitations of the MASON random-number generator, I use a uniform approximation \citep{weisstein_2013}, defined as:
+$$
+\left((x_{1}^{\alpha} - x_{0}^{\alpha})y + x_{0}^{\alpha}\right)^{\frac{1}{\alpha}} \sim \text{Power Law}
+$$
+where 
+$$r.v. y \sim U[0,1]$$
+$$ x_0  \equiv \text{minimum}$$
+$$ x_1  \equiv \text{maximum}$$
+$$ \alpha \equiv \text{Coefficient}$$
+
 \end{itemize}
@@
 \begin{itemize}
+	\item \textbf{Spare Capacity} is the country's capacity to rapidly increase oil for export above its inital baseline.
+Defaults to 0.1 (that is, countries can rapidly increase output by 10\%) based on estimation from \citet{mearns_2012}, \citet{iea_opec} and  consultations with subject-matter experts.
 	\item \textbf{Increasing Production} is a boolean of whether or not the country is currently in a state of temporarily-increased production.
@@
 When a country enters crisis, this number is drawn from a power law distribution, as described below.
-			\item \textbf{Total Capacity} is the country's capacity to rapidly increase oil for export above its inital baseline.
-Defaults to 1.1 (that is, countries can rapidly increase output by 10\%) based on consultations with subject-matter experts.
+
 \end{itemize}

### 2013-12-13 Writeups/CSS739 Paper/Paper.tex
Adding analysis code and near-final paper

@@
 	\usepackage{algorithm2e}
+	\usepackage{graphicx}
+	\usepackage{subcaption}
+	\usepackage{setspace}
+
+% set algorithm captions above psuedocode
+\makeatletter
+\renewcommand{\@algocf@capt@plain}{above} 
+\makeatother
 
@@
 	\author{David Masad}
+	\date{}
 	\maketitle
+
+\begin{abstract}
+Energy security is placed at risk by exogenous supply shocks, in particular political crises and conflicts that disrupt resource extraction and transportation.
+In this paper, a computational model of the security of international crude oil supplies is described, and its output analyzed.
+The model consists of country agents, linked geographically and by a directed, data-derived oil trade network.
+Countries stochastically experience crises, with probabilities and durations drawn randomly from data-driven distributions.
+The effect of these crises on secure oil supplies is measured globally and by country, and the effect of conflict contagion and spare production capacity are also estimated.
+The model indicates that Russia, Eastern Europe, and much of the Global South are at the greatest risk of supply shocks, while American producers are at greatest risk of demand shocks.
+It estimates that conflict contagion decreases energy security slightly, while spare capacity has minimal effect.
+\end{abstract}
+
+\onehalfspacing
 
@@
 Much of the analysis to date has taken a portfolio-based approach.
-\citet{wu_2009} uses portfolio diversity to analyze China's mix of crude oil and processed petrolem products.
+\citet{wu_2009} uses portfolio diversity to analyze China's mix of crude oil and processed petroleum products.
 The paper notes that ``Oil import-specific risk derives from events, such as political events, natural disaster and so on, that are not related to general movements in the international market'', but does not analyze the risks of such events directly.
@@
 However, the greater the diversity of suppliers (another input into the model), the greater the potential range of political stabilities.
-Averaging ratings of political risk eliminates the complexity that characterises international political stabilitiy, ignoring the possibility of crisis and conflict contagion.
+Averaging ratings of political risk eliminates the complexity that characterizes international political systems \citep{cederman_1997,geller_2011}, and ignores the possibility of crisis and conflict contagion \citep{black_2013}, by which conflict in one place can trigger other conflicts nearby.
 
 In fact, as \citet{hamilton_2003} describes, oil supply shocks have been driven caused by conflict involving oil producers, both domestic and international.
-\citet{hamilton_2003} argues that supply shock-driven changes to oil prices have a nonlinear effect on GDPs, while \citet{killian_2008} argues that the macroeconomic effects of the oil shocks themselves are small.
+\citet{hamilton_2003} argues that supply shock-driven changes to oil prices have a nonlinear effect on GDP, while \citet{kilian_2008} argues that the macroeconomic effects of the oil shocks themselves are small.
 Nevertheless, the price of oil is only an indirect indicator of energy security; if a substantial volume of oil cannot be extracted or exported, and hence is unavailable for consumption, it poses an energy security risk regardless of what the price of oil is doing.
 
-In this paper, I present a novel model linking energy security with political conflict.
+In this paper, I present a novel model linking energy security with political instability.
 At the core of the model is the supply and demand for crude oil at the country level, both imported and domestically-produced.
@@
 The model operates on the meso timescale, of months and years rather than days or decades.
-It produces local and global time-series of oil security indicators with each iteration.
+It produces country-level and global time-series of oil security indicators with each iteration.
 By running multiple iterations with different parameters, I generate notional near-future scenarios of changes in international energy security.
 
-The model straddles the boundary between a complicated cellular automata and a simple agent-based model.
-As we shall see, at its simplest configuration, the model is essentially stochastically sampling from the space of possible international crisis configurations and measuring their impact, with no agent decisionmaking or interaction.
-By incrementally adding more  elements (in particular, the possibility of conflict contagion and rapid increases of production to counterbalance supply shocks), we can estimate the impact of the additional complexity.
+The model is designed to be `as simple as possible, but no simpler.'  As such, it straddles the boundary between a complicated cellular automata on a network and a simple agent-based model.
+As we shall see, at its simplest configuration, the model is essentially stochastically sampling from the space of possible international crisis configurations and measuring their impact, with no agent decisionmaking.
+By incrementally adding elements (in particular, the possibility of conflict contagion and rapid increases of production to counterbalance supply shocks), we can estimate the impact of the additional complexity.
 
@@
 I use the model to conduct several experiments, describe the output, and conduct a sensitivity analysis.
-Finally, I discuss my conclucions, both regarding international energy security and the challenges and opportunities of this type of modeling.
+Finally, I discuss my conclusions, both regarding international energy security and the challenges and opportunities of this type of modeling.
 
@@
 The network of oil imports and exports comes from the United Nations-maintained COMTRADE database \citep{un_2013}.
-I extract all imports of crude petroleum (product code 2709) and other unprocessed petrolem (product code 2710) for 2012, the most recent year for which full data is available.
+I extract all imports of crude petroleum (product code 2709) and other unprocessed petroleum (product code 2710) for 2012, the most recent year for which complete data is available.
 Note that import data appears to be more reliable than export data: many major oil producers (particularly Saudi Arabia and Venezuela) only report export volume by region (e.g.
@@
 For the majority of countries, I assume that supply and demand are equal to total export and imports, respectively.
-For the top ten oil producers, I use US Energy Information Administration (EIA) estimates \citep{eia_2013} for those countries' consumption of their own domestic production.
+For the top ten oil producers, I use US Energy Information Administration (EIA) estimates \citep{eia_2013}, country estimates \citep{canada_2009,eia_domestic}, expert reports \citep{mohamedi_2010}, and media sources \citep{rasmi_2013} for those countries' consumption of their own domestic production.
+
+Spare capacity is particularly difficult to estimate, and varies over time based on the price of oil and other factors \citep{mearns_2012}.
+I use media sources such as \citet{daya_2012} and subject-matter expert (SME) analysis such as \citet{mearns_2012} in order to estimate the spare capacity of other countries, when possible.
+I also assign a default spare capacity of 10\% based on SME input.
 
@@
 These are in turn estimated based on the methodology developed by the Political Instability Task Force \citep{goldstone_2005}, which could predict the onset of internal instability in a two-year period with 80\% accuracy.
-Thus, these scores are normalized such that a rating of 10 (the maximum) is associated with an 80\% probability of instability within a 24-month timeframe, translating into a [[FILLIN]] monthly probability of crisis onset.
+Thus, these scores are normalized such that a rating of 10 (the maximum) is associated with an approximately 80\% probability of instability within a 24-month timeframe, translating into a 6.5\% monthly probability of crisis onset.
 
@@
 \citet{cioffi_2004} and others have argued that internal and external conflict durations follow a power law distribution, which I use here.
-I calibrate the parameters from two datasets: the UCDP/PRIO Armed Conflict Dataset \citep{lotta_2013} for military conflicts, and the Social Conflict in Africa Database \citep{hendrix_2013}, which includes both armed and lower-level social conflicts.
+I calibrate the power law coefficient from two datasets: the UCDP/PRIO Armed Conflict Dataset \citep{lotta_2013} for military conflicts, and the Social Conflict in Africa Database \citep{hendrix_2013}, which includes both armed and lower-level social conflicts.
 I subset the PRIO dataset for conflicts in Africa and merge it with SCAD, eliminating duplicates and events below the severity of a general strike, in order to obtain as complete as possible a set of crises and conflicts that may be associated with an oil production  shock.
-I find the duration of each event, and fit a power law to the resulting distribution, resulting in an estimated coefficient of $\mathbf{-1.37}$.
+I find the duration of each event in months, and fit a power law to the resulting distribution, resulting in an estimated coefficient of $\mathbf{-1.37}$.
 
@@
 
-The model consists of \textbf{Country} agents, posessing several characteristics and linked by two networks: a directed \textbf{trade network}, and an undirected \textbf{geographic adjacency} network, where the country agents are the network nodes.
+The model consists of \textbf{Country} agents, possessing several characteristics and linked by two networks: a directed \textbf{trade network}, and an undirected \textbf{geographic adjacency} network, where the country agents are the network nodes.
 
 Model variables may be \emph{fixed} based on data, \emph{parameters} that are fixed for each run and may be changed between them, or \emph{states} of the agents and edges that vary during the model.
+
+The model was written in Java, using the MASON package \citep{luke_2005}, with additional analysis done in Python.
 
@@
 It is obtained from \citet{eiu_2013}.
-	\item \textbf{In Crisis?} is simply a boolean variable determining whether or not a country is currently experiencing a crisis or conflict.
+	\item \textbf{In Crisis} is simply a boolean variable determining whether or not a country is currently experiencing a crisis or conflict.
 	\item \textbf{Crisis Length} is the number of months remaining in a crisis, if one is currently in process.
 When a country enters crisis, this number is drawn from a power law distribution.
-Specifically, in order to accomodate the limitations of the MASON random-number generator, I use a uniform approximation \citep{weisstein_2013}, defined as:
+Specifically, in order to accommodate the limitations of the MASON random-number generator, I use a uniform approximation \citep{weisstein_2013}, defined as:
 $$
@@
 \begin{itemize}
-	\item \textbf{Spare Capacity} is the country's capacity to rapidly increase oil for export above its inital baseline.
-Defaults to 0.1 (that is, countries can rapidly increase output by 10\%) based on estimation from \citet{mearns_2012}, \citet{iea_opec} and  consultations with subject-matter experts.
+	\item \textbf{Spare Capacity} is the country's capacity to rapidly increase oil for export above its initial baseline.
+Defaults to 0.1 (that is, countries can rapidly increase output by 10\%) based on estimation from \citet{mearns_2012,eia_opec} and  consultations with subject-matter experts.
 	\item \textbf{Increasing Production} is a boolean of whether or not the country is currently in a state of temporarily-increased production.
-	\item \textbf{Increasing Production For} stores the country being assisted. I assume that countries target their assistance at one other country.
-	\item \textbf{Crisis Length} is the number of months remaining in a crisis, if one is currently in process.
-When a country enters crisis, this number is drawn from a power law distribution, as described below.
+	\item \textbf{Increasing Production For} stores the country being assisted. I assume that countries target their assistance at one other country at any given time.
 
@@
 
-Each tick of the model, agents are activated in random order, as follows:
+Each tick of the model, agents are activated in random order. Each agent activation goes as follows:
 
@@
 
-I run the model for 250 iterations for each permutation of the Contagion and Assistance flags, 1,000 iterations in total; each iteration was run for 60 ticks, or 5 simulated years.
+I run the model for 250 iterations for each permutation of the Contagion and Assistance parameters, 1,000 iterations in total; each iteration was run for 60 ticks, or 5 simulated years.
 I collect the country-level and global indicators for each month of each iteration.
 
-Figure X shows a histogram of the Global Overall Ratio (the ratio between secure demand and supply) across all iteration-months.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\textwidth]{../Graphics/OverallDistribution}
+	\caption{Distribution of Overall Ratios across all scenario-months}
+
+\end{figure}
+
+Figure 1 shows a histogram of the Global Overall Ratio (the ratio between secure demand and supply) across all iteration-months.
 It appears to approximate a skew normal distribution, centered very close to 1 (a balance between supply and demand) but with a longer tail to the right -- supply shocks are more likely than demand shocks, and the most extreme supply shocks the model generates, though low-probability, are more extreme than any demand shocks.
 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\textwidth]{../Graphics/OverallVarMeanScatter}
+	\caption{Variance and mean of Overall Ratio for each model iteration}
+\end{figure}
+
 We can also examine the model iterations separately.
-Initially, we can characterize each iteration by the mean and variance of each indicator, as shown in Figure X.
+Initially, we can characterize each iteration by the mean and variance of each indicator, as shown in Figure 2.
+Note that the overall correlation between mean and variance is 0.33, indicating that they are not necessarily linearly related to one another.
 A high Overall or Supply mean ratio may indicate overall higher oil insecurity, but if the variance is low, it is `stably' insecure; in contrast, the mean may be close to 1 or even below it, but high variance would indicate severe uncertainty and fluctuation -- a distinct type of insecurity.
-Note that the overall correlation between mean and variance is XX, indicating that they are not directly linked to one another.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\textwidth]{../Graphics/AllOverallTraces}
+	\caption{Traces of Overall Ratio for all 1,000 iterations}
+\end{figure}
 
 We can further look at the trace of the indicators within each iteration, as they change from tick to tick.
-Figure X shows the traces of the 1,000 model runs overlaid on one another, and indicates some clear patterns.
+Figure 3 shows the traces of the 1,000 model runs overlaid on one another, and indicates some clear patterns.
 The majority of runs stay within a fairly narrow band, indicating similar levels of stability.
 Spikes outside the band, when they occur, tend to be brief.
-Note, however, that there are some longer-term spikes -- shocks which persist for multiple months, yielding values that become at least temporary baselines, even seeing fluctuations around them.
+Note, however, that there are some longer-term plateaus and troughs -- shocks which persist for multiple months, yielding values that become at least temporary baselines, even seeing fluctuations around them.
 
@@
 In order to disaggregate the effects of the Contagion and Assistance model parameters, we examine their results separately.
-Figure XX shows histograms of the Overall Ratio for the scenario-months for Contagion and Assistance, respectively.
+Figure 4 shows probability densities of the Overall Ratio for the scenario-months for Contagion and Assistance, respectively.
 
-Upon visual examination, the distributions with Contagion set to `On' and `Off' appear extremely similar.
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}{\textwidth}
+		\includegraphics[width=\textwidth]{../Graphics/ContagionDistribution}
+		\caption{Effect of Contagion}
+	\end{subfigure}
+
+	\begin{subfigure}{\textwidth}
+		\includegraphics[width=\textwidth]{../Graphics/AssistanceDistribution}
+		\caption{Effect of Assistance}
+	\end{subfigure}
+
+	\caption{Distribution of Overall Ratios with model parameters turned on and off}
+\end{figure}
+
+Upon visual examination, the distributions with Contagion enabled and disabled appear extremely similar.
 I conduct several statistical tests to determine whether they are in fact drawn from the same underlying distribution -- that is, whether there is no actual difference between them.
@@
 
-Figure XX shows a map of the overall risk of a major supply crisis, with the supply ratio rising above 1.5 (that is, demand exceeds secure supply by at least 50\%).
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\textwidth]{../Graphics/SupplyShockMap}
+	\caption{Risk of a major supply shock (Supply Ratio $>2$)}
+
+\end{figure}
+
+Figure 5 shows a map of the overall risk of a major supply crisis, with the supply ratio rising above 2 (that is, demand exceeds secure supply by at least 100\%).
 Several things immediately jump out: major oil producers appear to be at the least risk, as their consumption of domestic production reduces their exposure to import disruption.
@@
 
-Figure XX shows a similar map for demand risk, and indicates the likelihood of total possible exports exceed secure exports by at least 50\%.
-While there are fewer countires with very low risk, there are fewer with extremely high risk as well.
-Interestingly, the major Middle East producers appear to face relatively low demand risk, likely indicating a robust set of trade relationships with realively stable countries.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\textwidth]{../Graphics/DemandShockMap}
+	\caption{Risk of a major demand shock (Demand Ratio $<0.5$)}
+
+\end{figure}
+
+Figure 6 shows a similar map for demand risk, and indicates the likelihood of total possible exports exceed secure exports by at least 100\%.
+While there are fewer countries with very low risk, there are fewer with extremely high risk as well.
+Interestingly, the major Middle East producers appear to face relatively low demand risk, likely indicating a robust set of trade relationships with relatively stable countries.
 In contrast, the major non-US producers in the Americas -- Canada, Mexico and Venezuela -- appear to face higher demand risk.
@@
 The space of parameters and outputs to search is extremely large; for the purpose of this analysis, I will limit myself to two parameters: the exponential coefficient on the conflict length power law (calibrated from data, as described above) and the default spare production capacity, estimated from SME input.
-I will only examine the global ratios, as proxies for the behavior of the overall model.
 
@@
 
-Figure XX shows the mean and variance of the Global Ratios produced for each iteration-tick at each combination of the parameters.
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}{\textwidth}
+		\includegraphics[width=\textwidth]{../Graphics/SensitivityAnalysisOverall}
+		\caption{Overall Ratio}
+	\end{subfigure}
+
+	\begin{subfigure}{\textwidth}
+		\includegraphics[width=\textwidth]{../Graphics/SensitivityAnalysisSupply}
+		\caption{Supply Ratio}
+	\end{subfigure}
+
+	\caption{Sensitivity analysis of ratio means and variances}
+\end{figure}
+
+Figure 7 shows the mean and variance of the Global Overall Supply and Demand Ratios produced for each iteration-tick at each combination of the parameters.
 Visual examination shows that there is not significant change in the Overall Ratio; the range of outputs is relatively small, and no combination of inputs produces significant outliers.
@@
 As the conflict coefficient increases (reducing the probably of long conflicts), the variance is reduced as well.
-However, for the Supply and Demand Ratios, there is a clear change driven by conflict coefficient: as the probability of long conflicts increases, the mean of the Supply Ratio goes up and of the Demand Ratio goes down.
+However, for the Supply Ratio, there is a clear change driven by conflict coefficient: as the probability of long conflicts increases, the mean and variance increase as well (while the mean Demand Ratio, not shown, decreases).
 Note, however, that that there does not appear to be a clear critical point; the change relatively slow and linear.
@@
 Nevertheless, it shows that such extreme shocks may occur within the current configuration of the international oil system, and provides an estimate for their probability.
-The model provides evidence that consumption of domestic oil production provides a significant improvement in energy security, even in the absence of total energy independence.
+The model indicates that consumption of domestic oil production provides a significant improvement in energy security, even in the absence of total energy independence.
 It also highlights the importance of diverse import and export partners.
@@
 Analysis of the Contagion and Assistance parameters yields additional interesting insight.
-The effect of contagion is statistically significant but small -- perhaps suggesting the difficulty facing theongoing debate over the phenomenon in political science, but also that the debate's conclusions may not be very important to policy.
+The effect of contagion is statistically significant but small -- perhaps suggesting the difficulty facing the ongoing debate over the phenomenon in political science, but also that the debate's conclusions may not be very important to policy.
 More counterintuitively, perhaps, the Assistance parameter does not have a significant, measurable effect.
@@
 The interaction between agents is minimal and largely indirect, and with both Contagion and Assistance turned off, the model is essentially stochastically sampling from the space of possible crises and durations.
-Nevertheless, it is sufficient to capture the core aspects of the international oil system and produces useful results.
+Nevertheless, it is sufficient to capture the core aspects of the international oil system and produces useful results, demonstrating that the approach is viable.
+Future work can build on this framework, adding in additional submodels, dynamics and behaviors and measuring their effect against the baseline.
+Such additions may include explicit oil transportation and refining accounting for systemic choke-points; allowing countries to dynamically change their trading partners, supply and demand; incorporating the risks of non-political disasters both natural (e.g.
+Hurricane Katrina) and manmade (such as the Deepwater Horizon spill); and more.
 
-However, even such a simple model had extensive input requirements, with data coming from various sources of varying completeness and reliability.
-Expanding the model (for example, by adding additional energy sources, or incorporating oil refineries and processed petroleum) would significantly increase the data requirements.
+However, as we have seen, even the simple model presented here had extensive input requirements, with data coming from various sources of varying completeness and reliability.
+Expanding the model would significantly increase the data requirements.
 Adding dynamics and behavior pose an even greater challenge.
+There is no complete theory of states' energy decisionmaking -- in fact, such decisionmaking is almost certainly embedded within a broader policy framework, including both external interests and internal competition.
+It may be the case that adding additional elements will make the model more brittle without adding validity or increasing predictive power.
 
+Finally, the model presented here may have another application, as a decision-support tool for policymakers.
+The model's user interface (both via input files and a GUI) allows specific scenarios to be initiated and run, and their consequences and outputs evaluated.
+This may range from simply testing the effect of a particular events (e.g.
+a 12-month internal crisis in Saudi Arabia) to estimating the broader consequences of structural change (what would be the consequences if China were to become 10\% less stable, or if Iran were to begin to export to Europe).
+By quickly setting up and observing a `state of the world', a policymaker can augment her intuition, test hypotheses, and identify interesting issues or outliers for further qualitative or quantitative analysis.
+
+\singlespacing
 \bibliographystyle{chicago}

### 2013-12-13 Writeups/CSS739 Paper/Paper.tex
Update

@@
 The model inputs are empirical whenever possible.
-Data inputs are: oil trade relationships, country political instability, and country consumption of domestic oil production.
-Total supply and demand are derived largely from input data.
+Data inputs are oil trade relationships, country political instability,  country consumption of domestic oil production, and spare production capacity.
 
@@
 However, for the Supply Ratio, there is a clear change driven by conflict coefficient: as the probability of long conflicts increases, the mean and variance increase as well (while the mean Demand Ratio, not shown, decreases).
-Note, however, that that there does not appear to be a clear critical point; the change relatively slow and linear.
+Note, however, that that there does not appear to be a clear critical point; the change is relatively linear, and the outcomes all fall within a narrow band.
 Note also that the calibrated value of the parameter is in the higher-variance, higher-risk zone.
@@
 Adding dynamics and behavior pose an even greater challenge.
-There is no complete theory of states' energy decisionmaking -- in fact, such decisionmaking is almost certainly embedded within a broader policy framework, including both external interests and internal competition.
+There is no complete theory of states' energy decisionmaking -- in fact, such decisionmaking is almost certainly embedded within a broader policy framework, tied to both external interests and internal issues.
 It may be the case that adding additional elements will make the model more brittle without adding validity or increasing predictive power.

### 2014-01-19 Writeups/SBP Paper/Paper.tex
Adding SBP conference paper

--- 2013-12-13 Writeups/CSS739 Paper/Paper.tex
@@
-\documentclass{article}
-	\usepackage{natbib}
+\documentclass{llncs}
 	\usepackage{mathtools}
@@
 	\usepackage{graphicx}
-	\usepackage{subcaption}
 	\usepackage{setspace}
@@
 
-\onehalfspacing
-
 \section{Introduction}
 
-Energy security is defined by the International Energy Agency as ``the uninterrupted availability of energy sources at an affordable price''\citep{iea_2013}.
-As the world's energy demand continues to expand in the face of growing scarcity, energy security is becoming increasingly critical not just to developed countries but to developing ones as well \citep{yergin_2006}.
+Energy security is defined by the International Energy Agency as ``the uninterrupted availability of energy sources at an affordable price''\cite{iea_2013}.
+As the world's energy demand continues to expand in the face of growing scarcity, energy security is becoming increasingly critical not just to developed countries but to developing ones as well \cite{yergin_2006}.
 Thus, assessing the energy security of different countries is important in order to understand the overall state of the international system.
@@
 Much of the analysis to date has taken a portfolio-based approach.
-\citet{wu_2009} uses portfolio diversity to analyze China's mix of crude oil and processed petroleum products.
+\cite{wu_2009} uses portfolio diversity to analyze China's mix of crude oil and processed petroleum products.
 The paper notes that ``Oil import-specific risk derives from events, such as political events, natural disaster and so on, that are not related to general movements in the international market'', but does not analyze the risks of such events directly.
 It does, however, address the relatively less significant risk of piracy along several shipping routes.
-\citet{stirling_2010} and \citet{skea_2010} take a broader, more theoretical approach, and develops a multicriteria methodology for evaluating the diversity of an entire energy portfolio as it contributes to energy security.
+\cite{stirling_2010} and \cite{skea_2010} take a broader, more theoretical approach, and develops a multicriteria methodology for evaluating the diversity of an entire energy portfolio as it contributes to energy security.
 However, changing the composition of an entire energy portfolio takes a great deal of time -- it is a long-term process, measured in years or decades.
-Similarly, \citet{jacobson_2009} evaluates a set of possible future courses of action to jointly address climate, environmental and energy-security concerns -- again, courses of actions only implementable over the long term.
+Similarly, \cite{jacobson_2009} evaluates a set of possible future courses of action to jointly address climate, environmental and energy-security concerns -- again, courses of actions only implementable over the long term.
 
@@
 However, the greater the diversity of suppliers (another input into the model), the greater the potential range of political stabilities.
-Averaging ratings of political risk eliminates the complexity that characterizes international political systems \citep{cederman_1997,geller_2011}, and ignores the possibility of crisis and conflict contagion \citep{black_2013}, by which conflict in one place can trigger other conflicts nearby.
+Averaging ratings of political risk eliminates the complexity that characterizes international political systems \cite{cederman_1997,geller_2011}, and ignores the possibility of crisis and conflict contagion \cite{black_2013}, by which conflict in one place can trigger other conflicts nearby.
 
-In fact, as \citet{hamilton_2003} describes, oil supply shocks have been driven caused by conflict involving oil producers, both domestic and international.
-\citet{hamilton_2003} argues that supply shock-driven changes to oil prices have a nonlinear effect on GDP, while \citet{kilian_2008} argues that the macroeconomic effects of the oil shocks themselves are small.
+In fact, as \cite{hamilton_2003} describes, oil supply shocks have been driven caused by conflict involving oil producers, both domestic and international.
+\cite{hamilton_2003} argues that supply shock-driven changes to oil prices have a nonlinear effect on GDP, while \cite{kilian_2008} argues that the macroeconomic effects of the oil shocks themselves are small.
 Nevertheless, the price of oil is only an indirect indicator of energy security; if a substantial volume of oil cannot be extracted or exported, and hence is unavailable for consumption, it poses an energy security risk regardless of what the price of oil is doing.
@@
 
-The network of oil imports and exports comes from the United Nations-maintained COMTRADE database \citep{un_2013}.
+The network of oil imports and exports comes from the United Nations-maintained COMTRADE database \cite{un_2013}.
 I extract all imports of crude petroleum (product code 2709) and other unprocessed petroleum (product code 2710) for 2012, the most recent year for which complete data is available.
@@
 For the majority of countries, I assume that supply and demand are equal to total export and imports, respectively.
-For the top ten oil producers, I use US Energy Information Administration (EIA) estimates \citep{eia_2013}, country estimates \citep{canada_2009,eia_domestic}, expert reports \citep{mohamedi_2010}, and media sources \citep{rasmi_2013} for those countries' consumption of their own domestic production.
+For the top ten oil producers, I use US Energy Information Administration (EIA) estimates \cite{eia_2013}, country estimates \cite{canada_2009,eia_domestic}, expert reports \cite{mohamedi_2010}, and media sources \cite{rasmi_2013} for those countries' consumption of their own domestic production.
 
-Spare capacity is particularly difficult to estimate, and varies over time based on the price of oil and other factors \citep{mearns_2012}.
-I use media sources such as \citet{daya_2012} and subject-matter expert (SME) analysis such as \citet{mearns_2012} in order to estimate the spare capacity of other countries, when possible.
+Spare capacity is particularly difficult to estimate, and varies over time based on the price of oil and other factors \cite{mearns_2012}.
+I use media sources such as \cite{daya_2012} and subject-matter expert (SME) analysis such as \cite{mearns_2012} in order to estimate the spare capacity of other countries, when possible.
 I also assign a default spare capacity of 10\% based on SME input.
 
-Political instability estimates are taken from the Economist Intelligence Unit's Political Index ratings \citep{eiu_2013}.
-These are in turn estimated based on the methodology developed by the Political Instability Task Force \citep{goldstone_2005}, which could predict the onset of internal instability in a two-year period with 80\% accuracy.
+Political instability estimates are taken from the Economist Intelligence Unit's Political Index ratings \cite{eiu_2013}.
+These are in turn estimated based on the methodology developed by the Political Instability Task Force \cite{goldstone_2005}, which could predict the onset of internal instability in a two-year period with 80\% accuracy.
 Thus, these scores are normalized such that a rating of 10 (the maximum) is associated with an approximately 80\% probability of instability within a 24-month timeframe, translating into a 6.5\% monthly probability of crisis onset.
@@
 Crisis duration is drawn independently from onset.
-\citet{cioffi_2004} and others have argued that internal and external conflict durations follow a power law distribution, which I use here.
-I calibrate the power law coefficient from two datasets: the UCDP/PRIO Armed Conflict Dataset \citep{lotta_2013} for military conflicts, and the Social Conflict in Africa Database \citep{hendrix_2013}, which includes both armed and lower-level social conflicts.
+\cite{cioffi_2004} and others have argued that internal and external conflict durations follow a power law distribution, which I use here.
+I calibrate the power law coefficient from two datasets: the UCDP/PRIO Armed Conflict Dataset \cite{lotta_2013} for military conflicts, and the Social Conflict in Africa Database \cite{hendrix_2013}, which includes both armed and lower-level social conflicts.
 I subset the PRIO dataset for conflicts in Africa and merge it with SCAD, eliminating duplicates and events below the severity of a general strike, in order to obtain as complete as possible a set of crises and conflicts that may be associated with an oil production  shock.
@@
 
-The model was written in Java, using the MASON package \citep{luke_2005}, with additional analysis done in Python.
+The model was written in Java, using the MASON package \cite{luke_2005}, with additional analysis done in Python.
 
@@
 	\item \textbf{Domestic Share} of the country's oil demand satisfied by domestic production.
-For most countries, this defaults to 0; for the top ten oil producers, it is set based on US Energy Information Administration data \citep{eia_2013}.
+For most countries, this defaults to 0; for the top ten oil producers, it is set based on US Energy Information Administration data \cite{eia_2013}.
 	\item \textbf{Total Imports} is the sum of the current volume of all oil imports. 
@@
 	\item \textbf{Instability} is a measure of the country's political instability on a 0-10 scale, and determines the probability of a crisis or conflict -- a production shock.
-It is obtained from \citet{eiu_2013}.
+It is obtained from \cite{eiu_2013}.
 	\item \textbf{In Crisis} is simply a boolean variable determining whether or not a country is currently experiencing a crisis or conflict.
@@
 When a country enters crisis, this number is drawn from a power law distribution.
-Specifically, in order to accommodate the limitations of the MASON random-number generator, I use a uniform approximation \citep{weisstein_2013}, defined as:
+Specifically, in order to accommodate the limitations of the MASON random-number generator, I use a uniform approximation \cite{weisstein_2013}, defined as:
 $$
@@
 $$
-where 
-$$r.v. y \sim U[0,1]$$
-$$ x_0  \equiv \text{minimum}$$
-$$ x_1  \equiv \text{maximum}$$
-$$ \alpha \equiv \text{Coefficient}$$
+where $r.v. y \sim U[0,1]$, $ x_0  \equiv \text{minimum}$, $ x_1  \equiv \text{maximum}$, and $ \alpha \equiv \text{Coefficient}$
 
@@
 	\item \textbf{Spare Capacity} is the country's capacity to rapidly increase oil for export above its initial baseline.
-Defaults to 0.1 (that is, countries can rapidly increase output by 10\%) based on estimation from \citet{mearns_2012,eia_opec} and  consultations with subject-matter experts.
+Defaults to 0.1 (that is, countries can rapidly increase output by 10\%) based on estimation from \cite{mearns_2012,eia_opec} and  consultations with subject-matter experts.
 	\item \textbf{Increasing Production} is a boolean of whether or not the country is currently in a state of temporarily-increased production.
@@
 
-\begin{figure}[h!]
-	\centering
-	\begin{subfigure}{\textwidth}
-		\includegraphics[width=\textwidth]{../Graphics/ContagionDistribution}
-		\caption{Effect of Contagion}
-	\end{subfigure}
-
-	\begin{subfigure}{\textwidth}
-		\includegraphics[width=\textwidth]{../Graphics/AssistanceDistribution}
-		\caption{Effect of Assistance}
-	\end{subfigure}
-
-	\caption{Distribution of Overall Ratios with model parameters turned on and off}
-\end{figure}
-
 Upon visual examination, the distributions with Contagion enabled and disabled appear extremely similar.
@@
 
-\begin{figure}[h!]
-	\centering
-	\begin{subfigure}{\textwidth}
-		\includegraphics[width=\textwidth]{../Graphics/SensitivityAnalysisOverall}
-		\caption{Overall Ratio}
-	\end{subfigure}
-
-	\begin{subfigure}{\textwidth}
-		\includegraphics[width=\textwidth]{../Graphics/SensitivityAnalysisSupply}
-		\caption{Supply Ratio}
-	\end{subfigure}
-
-	\caption{Sensitivity analysis of ratio means and variances}
-\end{figure}
-
 Figure 7 shows the mean and variance of the Global Overall Supply and Demand Ratios produced for each iteration-tick at each combination of the parameters.
@@
 
-\singlespacing
-\bibliographystyle{chicago}
+\bibliographystyle{splncs}
 \bibliography{energysecurity}

### 2014-01-20 Writeups/SBP Paper/Paper.tex
Updaring SBP paper
Update SBP paper and tweak some graphics to look better at a smaller
size

@@
 
-% set algorithm captions above psuedocode
 \makeatletter
@@
 Energy security is defined by the International Energy Agency as ``the uninterrupted availability of energy sources at an affordable price''\cite{iea_2013}.
-As the world's energy demand continues to expand in the face of growing scarcity, energy security is becoming increasingly critical not just to developed countries but to developing ones as well \cite{yergin_2006}.
+As the world's energy demand continues to expand, energy security is becoming increasingly critical not just to developed countries but to developing ones as well \cite{yergin_2006}.
 Thus, assessing the energy security of different countries is important in order to understand the overall state of the international system.
 Despite this, there have been few attempts to create formal models of energy security.
-
-Much of the analysis to date has taken a portfolio-based approach.
-\cite{wu_2009} uses portfolio diversity to analyze China's mix of crude oil and processed petroleum products.
-The paper notes that ``Oil import-specific risk derives from events, such as political events, natural disaster and so on, that are not related to general movements in the international market'', but does not analyze the risks of such events directly.
-It does, however, address the relatively less significant risk of piracy along several shipping routes.
-\cite{stirling_2010} and \cite{skea_2010} take a broader, more theoretical approach, and develops a multicriteria methodology for evaluating the diversity of an entire energy portfolio as it contributes to energy security.
-However, changing the composition of an entire energy portfolio takes a great deal of time -- it is a long-term process, measured in years or decades.
-Similarly, \cite{jacobson_2009} evaluates a set of possible future courses of action to jointly address climate, environmental and energy-security concerns -- again, courses of actions only implementable over the long term.
-
-Short-term energy security, in contrast, appears to have been little-addressed in the literature.
-However, an important short-term analysis, MOSES (Model of Short-term Energy Security), was created within the International Energy Agency (IEA), the intergovernmental organization of the world's major oil importers established in part to protect their energy security.
-As the name of the model indicates, MOSES focuses on short-term risks, which it defines as a timescale of days and weeks.
-Furthermore, MOSES assigns risk scores based on a broad but ultimately static set of indicators.
-In particular, the indicator used to measure political stability is the ``weighted average of political stability of suppliers''.
-However, the greater the diversity of suppliers (another input into the model), the greater the potential range of political stabilities.
-Averaging ratings of political risk eliminates the complexity that characterizes international political systems \cite{cederman_1997,geller_2011}, and ignores the possibility of crisis and conflict contagion \cite{black_2013}, by which conflict in one place can trigger other conflicts nearby.
-
-In fact, as \cite{hamilton_2003} describes, oil supply shocks have been driven caused by conflict involving oil producers, both domestic and international.
-\cite{hamilton_2003} argues that supply shock-driven changes to oil prices have a nonlinear effect on GDP, while \cite{kilian_2008} argues that the macroeconomic effects of the oil shocks themselves are small.
-Nevertheless, the price of oil is only an indirect indicator of energy security; if a substantial volume of oil cannot be extracted or exported, and hence is unavailable for consumption, it poses an energy security risk regardless of what the price of oil is doing.
+[[EXPAND]]
 
@@
 As we shall see, at its simplest configuration, the model is essentially stochastically sampling from the space of possible international crisis configurations and measuring their impact, with no agent decisionmaking.
-By incrementally adding elements (in particular, the possibility of conflict contagion and rapid increases of production to counterbalance supply shocks), we can estimate the impact of the additional complexity.
+By incrementally adding elements (in particular, the possibility of conflict contagion and rapid increases of production to counterbalance supply shocks), we can estimate the impact of additional complexity.
 
 In the remainder of the paper, I describe the model in more detail, its data sources, and its various submodels.
-I use the model to conduct several experiments, describe the output, and conduct a sensitivity analysis.
+I use the model to conduct several experiment and describe the output.
 Finally, I discuss my conclusions, both regarding international energy security and the challenges and opportunities of this type of modeling.
@@
 The network of oil imports and exports comes from the United Nations-maintained COMTRADE database \cite{un_2013}.
-I extract all imports of crude petroleum (product code 2709) and other unprocessed petroleum (product code 2710) for 2012, the most recent year for which complete data is available.
-Note that import data appears to be more reliable than export data: many major oil producers (particularly Saudi Arabia and Venezuela) only report export volume by region (e.g.
-Europe, North America); however, the majority of their trade partners report their imports from them.
+I extract all imports of crude and other unprocessed petroleum for 2012, the most recent year for which complete data is available.
 
@@
 Political instability estimates are taken from the Economist Intelligence Unit's Political Index ratings \cite{eiu_2013}.
-These are in turn estimated based on the methodology developed by the Political Instability Task Force \cite{goldstone_2005}, which could predict the onset of internal instability in a two-year period with 80\% accuracy.
+These are in turn estimated based on the methodology developed by the Political Instability Task Force \cite{goldstone_2005}, which predicts the onset of internal instability in a two-year period with approximately 80\% accuracy.
 Thus, these scores are normalized such that a rating of 10 (the maximum) is associated with an approximately 80\% probability of instability within a 24-month timeframe, translating into a 6.5\% monthly probability of crisis onset.
@@
 I calibrate the power law coefficient from two datasets: the UCDP/PRIO Armed Conflict Dataset \cite{lotta_2013} for military conflicts, and the Social Conflict in Africa Database \cite{hendrix_2013}, which includes both armed and lower-level social conflicts.
-I subset the PRIO dataset for conflicts in Africa and merge it with SCAD, eliminating duplicates and events below the severity of a general strike, in order to obtain as complete as possible a set of crises and conflicts that may be associated with an oil production  shock.
+I subset and combine both datasets to generate as complete as possible a set of crises and conflicts that may be associated with oil production shocks in Africa.
 I find the duration of each event in months, and fit a power law to the resulting distribution, resulting in an estimated coefficient of $\mathbf{-1.37}$.
-
-\subsection{Formal Description}
-
-The model consists of \textbf{Country} agents, possessing several characteristics and linked by two networks: a directed \textbf{trade network}, and an undirected \textbf{geographic adjacency} network, where the country agents are the network nodes.
-
-Model variables may be \emph{fixed} based on data, \emph{parameters} that are fixed for each run and may be changed between them, or \emph{states} of the agents and edges that vary during the model.
-
-The model was written in Java, using the MASON package \cite{luke_2005}, with additional analysis done in Python.
-
-\subsubsection{Networks}
-
-Agents are connected by two networks, which we can also express as matrices.
-Let $\mathbf{G}$ be the geographic adjacency matrix, such that $g_{i,j}=g_{j,i}=1$ when counties $i$ and $j$ are geographically contiguous or near-contiguous, and $0$ otherwise.
-
-Let $\mathbf{E}$ be the trade matrix, such that $e_{i,j}$ denotes the volume of oil exports from $i$ to $j$.
-Note that if the assistance submodel is turned on, the values of $\mathbf{E}$ may change over time.
-
-\subsubsection{Agent parameters and variables}
-
-Each agent represents an independent country.
-Agents are associated with the \textbf{geometry} of their position on a world map, which determines their neighbors.
-They are described by two submodels: an oil model, and a crisis model.
-
-\paragraph{Oil Submodel}
-\begin{itemize}
-	\item \textbf{Domestic Share} of the country's oil demand satisfied by domestic production.
-For most countries, this defaults to 0; for the top ten oil producers, it is set based on US Energy Information Administration data \cite{eia_2013}.
-	\item \textbf{Total Imports} is the sum of the current volume of all oil imports. 
-	\item \textbf{Total Demand} is the sum of current imported and domestic oil consumed.
-Since the EIA reports the share of total consumption that is domestically-sourced, demand is calculated as: $$\text{Total Demand}_i = \sum_j e_{j,i} \cdot (1 - \text{Domestic Share})^{-1}$$
-	With $Total Imports$ set at the initial baseline. If current imports increase (as explained below) then supply will exceed demand. 
-	\item \textbf{Total Exports} is the sum of current exported oil to all other countries.
-\end{itemize}
-
-\paragraph{Crisis Submodel}
-\begin{itemize}
-	\item \textbf{Instability} is a measure of the country's political instability on a 0-10 scale, and determines the probability of a crisis or conflict -- a production shock.
-It is obtained from \cite{eiu_2013}.
-	\item \textbf{In Crisis} is simply a boolean variable determining whether or not a country is currently experiencing a crisis or conflict.
-	\item \textbf{Crisis Length} is the number of months remaining in a crisis, if one is currently in process.
-When a country enters crisis, this number is drawn from a power law distribution.
-Specifically, in order to accommodate the limitations of the MASON random-number generator, I use a uniform approximation \cite{weisstein_2013}, defined as:
-$$
-\left((x_{1}^{\alpha} - x_{0}^{\alpha})y + x_{0}^{\alpha}\right)^{\frac{1}{\alpha}} \sim \text{Power Law}
-$$
-where $r.v. y \sim U[0,1]$, $ x_0  \equiv \text{minimum}$, $ x_1  \equiv \text{maximum}$, and $ \alpha \equiv \text{Coefficient}$
-
-\end{itemize}
-
-\paragraph{Assistance Submodel}
-\begin{itemize}
-	\item \textbf{Spare Capacity} is the country's capacity to rapidly increase oil for export above its initial baseline.
-Defaults to 0.1 (that is, countries can rapidly increase output by 10\%) based on estimation from \cite{mearns_2012,eia_opec} and  consultations with subject-matter experts.
-	\item \textbf{Increasing Production} is a boolean of whether or not the country is currently in a state of temporarily-increased production.
-	\item \textbf{Increasing Production For} stores the country being assisted. I assume that countries target their assistance at one other country at any given time.
-
-\end{itemize}
-
-\subsubsection{Agent Behavior}
-
-Each tick of the model, agents are activated in random order. Each agent activation goes as follows:
-
-\begin{algorithm}[H]
-	\caption{main loop}
-	\eIf{not in crisis} {
-			Call crisis submodel \;
-		}{
-			$Crisis Length = Crisis Length - 1$ \;
-			\If {Crisis Length == 0} { End crisis \; }
-		}
-	\If{Assistance turned on} {
-		Call assistance submodel \;
-	}	
-\end{algorithm}
-
-\begin{algorithm}[H]
-	\caption{crisis submodel}
-		enter crisis with $Pr(crisis) \propto instability$ \;
-		\If{entered crisis}{
-			Crisis Length $\mathtt{\sim}  \text{Power Law}$ \;
-			\If{Contagion turned on} {
-				\For{each neighboring country} {
-					Call neighbor's crisis submodel \;
-			}}}
-\end{algorithm}
-
-\begin{algorithm}[H]
-	\caption{assistance submodel}
-	\eIf{Currently assisting another country} {
-			\If{Assisted country no longer in supply shock} {
-				\tcp{Ending assistance}
-				\For{each export edge} {
-					edge volume = baseline edge volume;\
-				}
-			}
-		} {
-			\For{each out-neighbor} {
-				\If{Neighbor in supply shock} {
-					Assist neighbor with $Pr = \frac{\text{Exports to neighbor}}{\text{Total exports}}$
-				}
-			}
-			\If{Assisting neighbor} {
-				\tcp{Increasing output}
-				\For{each export edge} {
-					edge volume = edge volume * (1 + excess capacity)\;
-				}
-			}
-		}
-
-\end{algorithm}
-
-Once all agents have acted, their energy-security metrics are computed, as described below.
-
-\subsubsection{Energy Security Metrics}
-
-Let $\delta(i)$ be the \emph{Security indicator} function, defined as:
-\[
-\delta(i) = \begin{dcases*}
-	1 & when $i$ is not in crisis\\
-	0 & when $i$ is in crisis
-\end{dcases*}
-\]
-
-Country-level indicators are:
-
-\begin{itemize}
-	\item \textbf{Supply Ratio} is the fraction of the country's current demand being met by \emph{secure} imports -- that is, imports from countries that are not in crisis. Formally:
-	$$
-		\text{Supply Ratio}_i = \frac{\text{Total Demand}_i}{\sum_{j}e_{j,i}\delta(j)}
-	$$
-\item \textbf{Demand Ratio} is the fraction of the country's current exports going to countries that are not in crisis -- secure exports. 
-$$
-\text{Demand Ratio}_i = \frac{\sum_{j}e_{i,j}\delta(j)}{\sum_{j}e_{i,j}}
-$$
-\end{itemize}
-
-The model also computes similar global ratios:
-\begin{itemize}
-	\item \textbf{Global Supply Ratio} is the all country's current demand being met by secure imports:
-	$$
-		\text{Global Supply Ratio} = \frac{\sum_i\text{Total Demand}_i}{\sum_i\sum_{j}e_{j,i}\delta(j)}
-	$$
-	\item \textbf{Global Demand Ratio} is the fraction of worldwide exports going to countries that are not in crisis:
-	$$
-		\text{Global Demand Ratio} = \frac{\sum_i \text{Total Demand}_i\delta(i)}{\sum_i\sum_{j}e_{i,j}}
-	$$
-	\item \textbf{Global Overall Ratio} is the ratio between secure supply and secure demand:
-	$$
-		\text{Global Overall Ratio} = \frac{\sum_i \text{Total Demand}_i\delta(i)}{\sum_i\sum_{j}e_{j,i}\delta(j)}
-	$$
-\end{itemize}
 
@@
 I run the model for 250 iterations for each permutation of the Contagion and Assistance parameters, 1,000 iterations in total; each iteration was run for 60 ticks, or 5 simulated years.
-I collect the country-level and global indicators for each month of each iteration.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{../Graphics/OverallDistribution}
+	\includegraphics[width=0.75\textwidth]{../Graphics/OverallDistribution_lg}
 	\caption{Distribution of Overall Ratios across all scenario-months}
@@
 
-Figure 1 shows a histogram of the Global Overall Ratio (the ratio between secure demand and supply) across all iteration-months.
-It appears to approximate a skew normal distribution, centered very close to 1 (a balance between supply and demand) but with a longer tail to the right -- supply shocks are more likely than demand shocks, and the most extreme supply shocks the model generates, though low-probability, are more extreme than any demand shocks.
+Figure 1 shows a histogram of the ratio between total global secure demand and supply across all iteration-months.
+It appears to approximate a skew normal distribution, centered very close to 1 (a balance between supply and demand) but with a longer tail to the right; extreme supply shocks are more likely than extreme demand shocks.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{../Graphics/OverallVarMeanScatter}
+	\includegraphics[width=0.75\textwidth]{../Graphics/OverallVarMeanScatter_lg}
 	\caption{Variance and mean of Overall Ratio for each model iteration}
@@
 
-We can also examine the model iterations separately.
-Initially, we can characterize each iteration by the mean and variance of each indicator, as shown in Figure 2.
-Note that the overall correlation between mean and variance is 0.33, indicating that they are not necessarily linearly related to one another.
-A high Overall or Supply mean ratio may indicate overall higher oil insecurity, but if the variance is low, it is `stably' insecure; in contrast, the mean may be close to 1 or even below it, but high variance would indicate severe uncertainty and fluctuation -- a distinct type of insecurity.
+I characterize each model iteration by the mean and variance of the supply-demand ratio, as shown in Figure 2.
+Overall correlation between mean and variance is 0.33, indicating that they are not necessarily linearly related to one another.
+A high mean ratio may indicate overall higher oil insecurity, but if the variance is low, it is `stably' insecure; in contrast, the mean may be close to 1 or even below it, but high variance would indicate severe uncertainty and fluctuation -- a distinct type of insecurity.
+
+\subsubsection{Contagion and Assistance}
+
+In order to disaggregate the effects of the Contagion and Assistance model parameters, I examine their results separately.
+Figure 4 shows probability densities of the Overall Ratio for the scenario-months for Contagion and Assistance, respectively.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{../Graphics/AllOverallTraces}
-	\caption{Traces of Overall Ratio for all 1,000 iterations}
+	\includegraphics[width=\textwidth]{../Graphics/ContagionAndAssistance}
+	\caption{Contagion and assistance}
 \end{figure}
 
-We can further look at the trace of the indicators within each iteration, as they change from tick to tick.
-Figure 3 shows the traces of the 1,000 model runs overlaid on one another, and indicates some clear patterns.
-The majority of runs stay within a fairly narrow band, indicating similar levels of stability.
-Spikes outside the band, when they occur, tend to be brief.
-Note, however, that there are some longer-term plateaus and troughs -- shocks which persist for multiple months, yielding values that become at least temporary baselines, even seeing fluctuations around them.
-
-\subsubsection{Contagion and Assistance}
-
-In order to disaggregate the effects of the Contagion and Assistance model parameters, we examine their results separately.
-Figure 4 shows probability densities of the Overall Ratio for the scenario-months for Contagion and Assistance, respectively.
-
 Upon visual examination, the distributions with Contagion enabled and disabled appear extremely similar.
-I conduct several statistical tests to determine whether they are in fact drawn from the same underlying distribution -- that is, whether there is no actual difference between them.
-Both the Kolmogorov-Smirnov  and Mann-Whitney tests indicate that the distributions are in fact different.
-Nevertheless, the differences are small.
-The mean ratio for the No Contagion scenario is 1.037, compared to 1.049 for the Contagion scenario.
-Both a parametric T-Test and nonparametric Wilcoxon Rank-Sum confirm that these means are indeed different, and neither the traditional (frequentist) not Bayesian confidence intervals overlap.
-Similarly, the variance for the No Contagion scenario is 0.025, while the Contagion variance is 0.03.
-Thus, conflict contagion pushes the scenarios away from the mean, yielding a wider range of outcomes.
-
+I conduct several statistical tests to determine whether they are in fact drawn from the same underlying distribution, and conclude that the difference is small but statistically significant.
+Conflict contagion pushes the scenarios away from the mean, yielding a wider range of outcomes.
 I repeat the same analysis for the the Assistance disabled and enabled scenarios.
 However, in this case, the statistical tests cannot reject the hypothesis that the resulting distributions are the same, and the means are within each other's confidence intervals.
-
-I conduct similar analyses for the other global ratios, for supply and demand, as well, which result in the same conclusion.
-The effects of Contagion are small but statistically significant, while Assistance has a far smaller and less significant effect.
 
@@
 
-\subsection{Sensitivity Analysis}
-
-As discussed above, the model accepts several input parameters, most of which are calibrated from data.
-Given that this calibration is only approximate, it is important to conduct a sensitivity analysis in order to identify whether the results are extremely sensitive to the input parameters -- the more sensitive they are, the more the model will be biased by missing or incomplete data, and the less reliable we expect it to be.
-The space of parameters and outputs to search is extremely large; for the purpose of this analysis, I will limit myself to two parameters: the exponential coefficient on the conflict length power law (calibrated from data, as described above) and the default spare production capacity, estimated from SME input.
-
-In order to conduct the sensitivity analysis, I iterate over conflict coefficient values between -1.05 and -2.0, and default spare production capacities between 0 and 0.5, both in steps of 0.05.
-For each combination, I run the model 40 times, 10 iterations each for each combination of Contagion and Assistance, for a total of 2,200 model runs.
-
-Figure 7 shows the mean and variance of the Global Overall Supply and Demand Ratios produced for each iteration-tick at each combination of the parameters.
-Visual examination shows that there is not significant change in the Overall Ratio; the range of outputs is relatively small, and no combination of inputs produces significant outliers.
-The one input-output combination detectable is between the conflict coefficient and the variance.
-As the conflict coefficient increases (reducing the probably of long conflicts), the variance is reduced as well.
-However, for the Supply Ratio, there is a clear change driven by conflict coefficient: as the probability of long conflicts increases, the mean and variance increase as well (while the mean Demand Ratio, not shown, decreases).
-Note, however, that that there does not appear to be a clear critical point; the change is relatively linear, and the outcomes all fall within a narrow band.
-Note also that the calibrated value of the parameter is in the higher-variance, higher-risk zone.
-In other words, the model may be forecasting risk on the higher end.
-
 \section{Discussion}
@@
 It also highlights the importance of diverse import and export partners.
-
 The model identifies the countries facing the greatest risks to their energy security, as well as making patterns in the risk visible -- particularly the high risk in the countries of the `Global South'.

### 2014-03-20 Writeups/SBP Paper/Paper.tex
Updated paper, began work on poster

@@
 	\usepackage{graphicx}
-	\usepackage{setspace}
+	\usepackage{hyperref}
 
@@
 	\author{David Masad}
-	\date{}
+	\institute{Department of Computational Social Science\\George Mason University\\Fairfax, VA}
 	\maketitle
@@
 In this paper, a computational model of the security of international crude oil supplies is described, and its output analyzed.
-The model consists of country agents, linked geographically and by a directed, data-derived oil trade network.
-Countries stochastically experience crises, with probabilities and durations drawn randomly from data-driven distributions.
+The model consists of country agents, linked geographically and by a data-derived oil trade network.
+Countries stochastically experience crises, with probabilities and durations drawn randomly from data-fitted distributions.
 The effect of these crises on secure oil supplies is measured globally and by country, and the effect of conflict contagion and spare production capacity are also estimated.
@@
 As the world's energy demand continues to expand, energy security is becoming increasingly critical not just to developed countries but to developing ones as well \cite{yergin_2006}.
-Thus, assessing the energy security of different countries is important in order to understand the overall state of the international system.
-Despite this, there have been few attempts to create formal models of energy security.
-[[EXPAND]]
+Much of the previous formal analysis of energy security has taken a portfolio-based approach, examining both the short-term risks\cite{wu_2009,jewell_2011} and long-term energy and environmental policy \cite{stirling_2010,skea_2010,jacobson_2009}.
+However, these approaches tend to aggregate political risk into a single indicator, and do not incorporate the complexity that characterizes international political in general, and conflict in particular \cite{cederman_1997,geller_2011}.
+In fact, oil supply shocks are often caused by domestic or international conflicts involving oil producers \cite{hamilton_2003}.
+While the macroeconomic effects of oil shocks may be small \cite{kilian_2008}, the price of oil is only an indirect measure of energy security.
+If a substantial volume of oil cannot be extracted or exported, and hence is unavailable for consumption, this poses an energy security risk regardless of the benchmark price.
 
 In this paper, I present a novel model linking energy security with political instability.
-At the core of the model is the supply and demand for crude oil at the country level, both imported and domestically-produced.
+At the core of the model is the supply and demand for crude oil at the country level.
 Supply and demand are placed at risk when countries experience crises and conflicts.
@@
 By running multiple iterations with different parameters, I generate notional near-future scenarios of changes in international energy security.
+By incrementally enabling the possibility of conflict contagion, and of rapid increases of production to counterbalance supply shocks, I can estimate their impact on international energy security.
 
-The model is designed to be `as simple as possible, but no simpler.'  As such, it straddles the boundary between a complicated cellular automata on a network and a simple agent-based model.
-As we shall see, at its simplest configuration, the model is essentially stochastically sampling from the space of possible international crisis configurations and measuring their impact, with no agent decisionmaking.
-By incrementally adding elements (in particular, the possibility of conflict contagion and rapid increases of production to counterbalance supply shocks), we can estimate the impact of additional complexity.
-
-In the remainder of the paper, I describe the model in more detail, its data sources, and its various submodels.
+In the remainder of the paper, I describe the model in more detail, as well as its input data sources.
 I use the model to conduct several experiment and describe the output.
@@
 
-\section{Model Description}
+\section{Model}
 
-\subsection{Overview}
+\subsection{Model Description}
 
-The model consists of countries, each of which has a set supply of and demand for crude oil.
-Countries are linked geographically, and by a network of export-import relationships.
-Each month, some countries enter crisis at random; the exports and imports of countries in crisis are considered \emph{at risk}.
-If \textbf{Contagion} is enabled, a crisis in one country may spread to its immediate geographic neighbors.
-If \textbf{Assistance} is enabled, oil exporters may increase their production to balance a loss of secure oil by their major import partners.
+The model consists of countries, each of which has a set supply of and demand for crude oil\footnote{Full model code is available at \url{https://github.com/dmasad/riskworld}}.
+Countries are linked geographically to their neighbors, and by a network of export-import relationships.
+Each monthly tick of the model, some countries enter crisis at random with probability based on their political instability.
+Crisis duration is randomly drawn from a power law distribution.
+The exports and imports of countries in crisis are considered \emph{at risk}.
+At the end of each tick, the model measures several indicators of oil security.
+The Overall Ratio is the global ratio of secure demand to secure supply; a high ratio indicates a higher likelihood of a global supply shock, while a low ratio indicates a higher likelihood of a global demand shock.
+For each country, the model computes the ratio of total demand to secure (not at risk) supply (the Supply Ratio).
+The higher this ratio is, the more likely the country is to experience a supply shock.
+The model also computes the fraction of each country's total exports are going to countries not in crisis (Demand Ratio), estimating the risk of a demand shock.
 
-The model parameters currently remain fixed for each run: political instability, supply, demand, and trade relations are treated as constant.
+If \textbf{Contagion} is enabled, when a country enters crisis it may trigger crises in neighboring countries based on their own instability.
+If \textbf{Assistance} is enabled, oil exporters may utilize their spare capacity and increase their production to balance a loss of secure oil by a major import partner, increasing the supply of oil to all of their trading partners.
 
 \subsection{Data Sources}
-The model inputs are empirical whenever possible.
-Data inputs are oil trade relationships, country political instability,  country consumption of domestic oil production, and spare production capacity.
-
 The network of oil imports and exports comes from the United Nations-maintained COMTRADE database \cite{un_2013}.
@@
 
-For the majority of countries, I assume that supply and demand are equal to total export and imports, respectively.
+For the majority of countries, I assume that supply and demand are equal to total exports and imports.
 For the top ten oil producers, I use US Energy Information Administration (EIA) estimates \cite{eia_2013}, country estimates \cite{canada_2009,eia_domestic}, expert reports \cite{mohamedi_2010}, and media sources \cite{rasmi_2013} for those countries' consumption of their own domestic production.
 
-Spare capacity is particularly difficult to estimate, and varies over time based on the price of oil and other factors \cite{mearns_2012}.
-I use media sources such as \cite{daya_2012} and subject-matter expert (SME) analysis such as \cite{mearns_2012} in order to estimate the spare capacity of other countries, when possible.
+Spare capacity is particularly difficult to estimate, and varies over time and with the price of oil and other factors \cite{mearns_2012}.
+I use media sources \cite{daya_2012} and subject-matter expert (SME) analysis \cite{mearns_2012} in order to estimate the spare capacity of major exporters.
 I also assign a default spare capacity of 10\% based on SME input.
@@
 
-Crisis duration is drawn independently from onset.
-\cite{cioffi_2004} and others have argued that internal and external conflict durations follow a power law distribution, which I use here.
+Crisis duration is drawn from a power-law distribution \cite{cioffi_2004}.
 I calibrate the power law coefficient from two datasets: the UCDP/PRIO Armed Conflict Dataset \cite{lotta_2013} for military conflicts, and the Social Conflict in Africa Database \cite{hendrix_2013}, which includes both armed and lower-level social conflicts.
@@
 
-\subsubsection{Outcome Range}
+\subsubsection{Outcome Range:}
 
@@
 	\centering
-	\includegraphics[width=0.75\textwidth]{../Graphics/OverallDistribution_lg}
-	\caption{Distribution of Overall Ratios across all scenario-months}
+	\includegraphics[width=0.75\textwidth]{Figures/OverallDistribution_lg}
+	\caption{Distribution of overall ratios across all scenario-months}
 
@@
 
-Figure 1 shows a histogram of the ratio between total global secure demand and supply across all iteration-months.
-It appears to approximate a skew normal distribution, centered very close to 1 (a balance between supply and demand) but with a longer tail to the right; extreme supply shocks are more likely than extreme demand shocks.
+Figure 1 shows a histogram of the ratio between total global secure demand and supply (overall ratio) across all iteration-months.
+It is centered very close to 1 (a balance between supply and demand) but with a longer tail to the right; extreme supply shocks are more likely than extreme demand shocks.
 
@@
 	\centering
-	\includegraphics[width=0.75\textwidth]{../Graphics/OverallVarMeanScatter_lg}
-	\caption{Variance and mean of Overall Ratio for each model iteration}
+	\includegraphics[width=0.75\textwidth]{Figures/OverallVarMeanScatter_lg}
+	\caption{Variance and mean of overall ratio for each model iteration}
 \end{figure}
 
-I characterize each model iteration by the mean and variance of the supply-demand ratio, as shown in Figure 2.
-Overall correlation between mean and variance is 0.33, indicating that they are not necessarily linearly related to one another.
-A high mean ratio may indicate overall higher oil insecurity, but if the variance is low, it is `stably' insecure; in contrast, the mean may be close to 1 or even below it, but high variance would indicate severe uncertainty and fluctuation -- a distinct type of insecurity.
+I characterize each model iteration by the mean and variance of the overall ratio, as shown in Figure 2.
+Correlation between mean and variance is 0.33, indicating that they are not necessarily linearly related to one another.
+A high mean ratio may indicate higher oil insecurity, but if the variance is low, it is `stably' insecure; in contrast, the mean may be close to 1 or even below it, but high variance would indicate severe uncertainty and fluctuation -- a distinct type of insecurity.
 
-\subsubsection{Contagion and Assistance}
+\subsubsection{Contagion and Assistance:}
 
 In order to disaggregate the effects of the Contagion and Assistance model parameters, I examine their results separately.
-Figure 4 shows probability densities of the Overall Ratio for the scenario-months for Contagion and Assistance, respectively.
+Figure 3 shows probability densities of the Overall Ratio for the scenario-months for Contagion and Assistance, respectively.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{../Graphics/ContagionAndAssistance}
-	\caption{Contagion and assistance}
+	\includegraphics[width=\textwidth]{Figures/ContagionAndAssistance}
+	\caption{Effects of Contagion and Assistance}
 \end{figure}
 
-Upon visual examination, the distributions with Contagion enabled and disabled appear extremely similar.
+The distributions with Contagion enabled and disabled appear extremely similar.
 I conduct several statistical tests to determine whether they are in fact drawn from the same underlying distribution, and conclude that the difference is small but statistically significant.
-Conflict contagion pushes the scenarios away from the mean, yielding a wider range of outcomes.
-I repeat the same analysis for the the Assistance disabled and enabled scenarios.
+Conflict contagion increases the mean and variance, yielding a wider range of outcomes.
+I repeat the same analysis for the Assistance parameter.
 However, in this case, the statistical tests cannot reject the hypothesis that the resulting distributions are the same, and the means are within each other's confidence intervals.
@@
 In addition to the aggregate global indicators, the model also tracks the Supply and Demand ratios for each country.
-We can analyze this output in order to understand how different countries' risks differ from one another, and to identify countries at high or low risk.
+We can analyze this output in order to understand how countries' risks differ from one another, and to identify countries at high or low risk.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{../Graphics/SupplyShockMap}
+	\includegraphics[width=\textwidth]{Figures/SupplyShockMap}
 	\caption{Risk of a major supply shock (Supply Ratio $>2$)}
@@
 
-Figure 5 shows a map of the overall risk of a major supply crisis, with the supply ratio rising above 2 (that is, demand exceeds secure supply by at least 100\%).
+Figure 4 shows a map of the overall risk of a major supply crisis, with the supply ratio rising above 2 (that is, demand exceeds secure supply by at least 100\%).
 Several things immediately jump out: major oil producers appear to be at the least risk, as their consumption of domestic production reduces their exposure to import disruption.
@@
 	\centering
-	\includegraphics[width=\textwidth]{../Graphics/DemandShockMap}
+	\includegraphics[width=\textwidth]{Figures/DemandShockMap}
 	\caption{Risk of a major demand shock (Demand Ratio $<0.5$)}
@@
 
-Figure 6 shows a similar map for demand risk, and indicates the likelihood of total possible exports exceed secure exports by at least 100\%.
+Figure 5 shows a similar map for demand risk, and indicates the likelihood of total possible exports exceeding secure exports by at least 100\%.
 While there are fewer countries with very low risk, there are fewer with extremely high risk as well.
@@
 The effect of contagion is statistically significant but small -- perhaps suggesting the difficulty facing the ongoing debate over the phenomenon in political science, but also that the debate's conclusions may not be very important to policy.
-More counterintuitively, perhaps, the Assistance parameter does not have a significant, measurable effect.
+More counterintuitively, perhaps, the Assistance parameter does not have a significant effect.
 Here is somewhere the model can generate insights that are not otherwise obvious -- specifically, that even if major oil producers can rapidly increase their output in response to crises, it is not enough to counterbalance the effects of those crises.
@@
 Finally, the model presented here may have another application, as a decision-support tool for policymakers.
-The model's user interface (both via input files and a GUI) allows specific scenarios to be initiated and run, and their consequences and outputs evaluated.
+The model's user interface (both via input files and a graphical user interface) allows specific scenarios to be initiated and run, and their consequences and outputs evaluated.
 This may range from simply testing the effect of a particular events (e.g.

### 2014-03-21 Analysis/ModelJSONParse.py
Adding high-rez graphics for the poster

@@ -1,0 +1,58 @@
+import pandas
+import json
+
+def make_dataframes(path, country_data=True):
+    '''
+    Load a model output JSON and parse it into two Pandas DataFrame, one for 
+    globals and another for country-level parameters.
+    '''
+
+    data = json.load(open(path))
+
+    # Build the variable lists
+    parameters = []
+    global_series = []
+    country_series = []
+    for var in data[0]:
+        if type(data[0][var]) is dict:
+            country_series.append(var)
+        elif type(data[0][var]) is list:
+            global_series.append(var)
+        else:
+            parameters.append(var)
+
+    # Build DataFrames
+    global_rows = []
+    country_rows = []
+    if country_data:
+        country_list = data[0][country_series[0]].keys()
+
+    for n, iteration in enumerate(data):
+        params = {p: iteration[p] for p in parameters}
+        params["Iteration"] = n
+        for i in range(60): # Hard-code number of ticks, for now
+            # Build global row
+            row = {"Tick": i}
+            for key in global_series:
+                row[key] = iteration[key][i]
+            row.update(params)
+            global_rows.append(row)
+            
+            if not country_data: continue
+
+            # Build local rows
+            for country in country_list:
+                row = {"Tick": i, "Country": country}
+                for key in country_series:
+                    row[key] = iteration[key][country][i]
+                row.update(params)
+                country_rows.append(row)
+
+    global_df = pandas.DataFrame(global_rows)
+    if country_data:
+        country_df = pandas.DataFrame(country_rows)
+        return (global_df, country_df)
+    else:
+        return global_df
+
+
